---
title: "Data preparation"
author: "Camille Crapart"
date: "2023-02-21"
output: 
   bookdown::html_document2:
     code_folding: hide
     toc: true
     toc_float: true
     number_sections: true
     fig_caption: true

bibliography: C:\\Users\\raine\\Documents\\UiO\\Bibtex\\fCO2.bib
link-citations: yes
---

html version: [Data preparation](https://camilmc.github.io/fCO2/Data_preparation.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = T, warning = F, error = F, fig.align = "center", results = T, collapse = T, cache.lazy = F)
options(knitr.kable.NA="", knitr.table.format = "html")
```

```{r libraries}
library(readxl)

library(AquaEnv)
library(MASS)
library(mgcv)

library(dplyr)
library(ggplot2)
library(colorspace)

library(sf)
library(sp)

fCO2.atm.2020 <- 411.51 * 1e-6 # oct 2020
fCO2.atm.2019 <- 408.75 * 1e-6 # oct 2019
fCO2.atm.2004 <- 374.63 * 1e-6 # oct 2004
fCO2.atm.1995 <- 360.17 * 1e-6 # 1995
```

Atmospheric concentration of CO2 is retrieved from: <https://climate.nasa.gov/vital-signs/carbon-dioxide/>, dataset on <https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.txt>


# Data from CBA and N112 surveys

TOC was analysed by infrared CO2 detection after catalytic high temperature combustion (Shimadzu TOC-VWP analyzer) at the Department of Biosciences at the University of Oslo. pH and alkalinity (end-point titration to pH 4.5) were measured at the Chemistry Department of the University of Oslo.  

## Data from CBA survey

The data is available at @CBA2019. Lake area and catchment area were provided by NIVA. 

```{r cba-data}
cba.all <- readxl::read_xlsx("CBA_100Lakes_Master.xlsx")

ggplot(cba.all)+geom_point(aes(x = TOC, y = as.character(Lake_ID)))+geom_vline(xintercept = 45)+labs(y="Lake ID")+theme_bw()
ggplot(cba.all)+geom_point(aes(x = Alkalinity, y = as.character(Lake_ID)))+geom_vline(xintercept = 0.7)+labs(y="Lake ID")+theme_bw()
ggplot(cba.all)+geom_point(aes(x = pH_Kje, y = as.character(Lake_ID)))+
  geom_vline(xintercept = 5)+labs(x = "pH", y="Lake ID")+theme_bw()

bdg.niva <- readRDS("bdg.niva.rds") %>% dplyr::select(c("CBA_Lake_ID", "lake_area_km2","basin_area_km2"))

cba <- merge(cba.all,bdg.niva, by.x = "Lake_ID", by.y = "CBA_Lake_ID") 
```

The wind speep raster was downloaded from @C3S2020. Given the resolution of the raster compared to the area of the studied lakes, we used the wind speed at 10 m from the surface at the coordinates of the sampling. (Resolution : 1x1 degree \~11.1x11.1 km \<- 123 km2)


```{r cba-windspeed, eval = F}

nws.files <- list.files(path = "NWS", pattern = "Wind_WFDE5_CRU_2019", full.names = T) 
nws.list.stacks <- lapply(nws.files, raster::stack) %>% lapply(raster::crop, c(0,35,55,73))
nws.list.means <- lapply(nws.list.stacks, raster::mean, na.rm = T)
saveRDS(nws.list.means, "cba.nws.list.mean.rds")
nws.2019 <- nws.list.means %>% raster::stack() %>% raster::mean()

cba.spdf <- SpatialPointsDataFrame(coords = cba[,c("Long","Lat")], data = cba[,c("Long","Lat","Lake_ID")], proj4string = CRS("+proj=longlat +datum=WGS84"))
annual.nws.cba <- raster::extract(nws.2019, cba.spdf, sp = T)

saveRDS(annual.nws.cba, "annual.nws.cba.rds")
```

The CBA dataset was then merged with the wind speed. 

```{r merge-cba}
annual.nws.cba <- readRDS("annual.nws.cba.rds")@data %>% select(c("Lake_ID","layer")) %>% setNames(c("Lake_ID","nws_m.s"))
cba <- cba %>%  merge(annual.nws.cba, by = "Lake_ID")
```

## N112 data

This survey was first published in @Larsen2011. 

```{r N112}
n112.all <- read.table("Larsen/N112_subset_110909.txt", sep="", header = TRUE,na.strings = c(".", "NA"))

n112 <- n112.all %>% subset(ykoord > 6e6) %>% subset(is.na(TIC.ug) == F) %>% subset(is.na(alkendpoint) == F)

n112.spdf <- SpatialPointsDataFrame(coords = n112[,c("xkoord","ykoord")], data = n112[,c("xkoord","ykoord")], proj4string = CRS("+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
119
 
n112.spdf.wgs <- n112.spdf %>% spTransform(CRS("+proj=longlat +datum=WGS84"))
```

Similarly, the wind speed was aquired from @C3S2020.

```{r n112-annual-wind-speed, eval = F}

nws.files <- list.files(path = "NWS", pattern = "Wind_WFDE5_CRU_2004", full.names = T) 
nws.list.stacks <- lapply(nws.files, raster::stack) %>% lapply(raster::crop, c(0,35,55,73))
nws.list.means <- lapply(nws.list.stacks, raster::mean, na.rm = T)
saveRDS(nws.list.means, "n112.nws.list.mean.rds")

nws.list.mean <- readRDS("n112.nws.list.mean.rds")
nws.2004 <- nws.list.mean %>% raster::stack() %>% raster::mean()

n112.spdf <- SpatialPointsDataFrame(coords = n112[,c("xkoord","ykoord")], data = n112[,c("xkoord","ykoord","Vatnlnr")], proj4string = CRS("+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

annual.nws.n112 <- raster::extract(nws.2004, n112.spdf, sp = T)

saveRDS(annual.nws.n112, "annual.nws.n112.rds")
```

Then the original dataset was merged with the wind speed. 

```{r merge-n112}
annual.nws.n112 <- readRDS("annual.nws.n112.rds")@data %>% dplyr::select(c("Vatnlnr","layer")) %>% setNames(c("Vatnlnr","nws_m.s"))

n112 <- merge(n112, annual.nws.n112, by = "Vatnlnr")
```

## Merge CBA and N112 data

Both dataset were merged in one. Concentration of TOC, TIC and other elements were converted in mol/L. 

```{r norway}
norway <- data.frame(survey = c(rep("CBA_2019", dim(cba)[1]),rep("N112_2004",dim(n112)[1])))

norway$lake.id <- c(cba$Lake_ID,
                    n112$Vatnlnr)
  
norway$lake.name <- c(cba$Lake_name,
                      rep(NA,dim(n112)[1]))

norway$TOC <- c(cba$TOC/12.011 * 1e-3, # mg/L to mol/L
                n112$TOC/12.011 * 1e-3) # mg/L to mol/L

norway$TIC <- c(rep(NA, dim(cba)[1]),
                n112$TIC.ug/12.011 * 1e-6) # ug/L to mol/L

norway$Ca <- c(cba$Ca / 40.78 * 1e-3, #mg/L to mol/L
                rep(NA, dim(n112)[1])) # ug/L to mol/L

norway$TN <- c(cba$TN / 14 * 1e-3, #mg/L to mol/L
                n112$TotN / 14 * 1e-6) # ug/L to mol/L

norway$Fe <- c(cba$Fe / 55.845 * 1e-3, #mg/L to mol/L
                rep(NA, dim(n112)[1]))

norway$TA <- c(cba$Alkalinity * 1e-3, # convert from meq/L to mol/L
               n112$alkendpoint * 1e-6) # assumes ueq/L to mol/L,

norway$temp_c <- c(cba$T,
                   n112$temp.situ)

norway$temp_k <- norway$temp_c + 273.15
  
norway$EC <- c(cba$EC_Kje,
               rep("NA",dim(n112)[1])) %>% as.numeric()

norway$IS <- 1.3e-5 * norway$EC

norway$fCO2.atm <- c(rep(fCO2.atm.2019, dim(cba)[1]), 
                     rep(rep(fCO2.atm.2004,dim(n112)[1])))

norway$pH <- c(cba$pH_Kje,
               n112$pH)

norway$Hplus <- 10^(-norway$pH)

norway$CO2 <- c(as.numeric(cba$CO2) * 1e-6, # from umol/L to mol/L
                n112$co2.umolar * 1e-6) # umol/L to mol/kg_solution
                
norway$long <- c(cba$Long,
                 n112.spdf.wgs@coords[,1])

norway$lat <- c(cba$Lat,
                 n112.spdf.wgs@coords[,2])

norway$lake.area <- c(cba$lake_area_km2,
                      n112$lake.area)

norway$catchment.area <- c(cba$basin_area_km2,
                      n112$poly.area)

norway$nws_m.s <- c(cba$nws_m.s,
                    n112$nws_m.s)

norway <- filter(norway, !is.na(TA))
```

We calculated the dissociation constants for the carbonate equilibrium based on @Dickson1979, @Harned1941, @Harned1943 and @Weiss1974.

```{r constants}

norway$Kw = with(norway, exp(148.9802 - 13847.26/temp_k - 23.6521*log(temp_k))) # Dickson and Riley, 1979

#norway$K0 <- with(norway, 10^-((-2622.38/temp_k) + 15.5873 - 0.0178471*temp_k)) #Harned and Davis, 1943
norway$K0 <- with(norway, exp(-58.0931+90.5069*100/temp_k+22.2940*log(temp_k/100))) # Weiss 1974
norway$K1 <- with(norway, 10^-((3404.71/temp_k)- 14.8435 + 0.032786*temp_k)) # Harned and Davis, 1943
norway$K2 <- with(norway, 10^-((2902.39/temp_k) - 6.4980 + 0.02379*temp_k)) # Harned and Scholes, 1941

```

## Carbonate speciation from measured CO2

Our datasets contain CO2 concentration calculated from total inorganic carbon, following the methods suggested by @Sobek2003 for the N112 survey and @Aberg2014 for the CBA survey.

The concentration of each carbonate species was calculated based on this value.

```{r carbonate-speciation-from-measured-CO2}
norway$H2CO3 <- norway$CO2 # in mol/L
norway$HCO3 <- with(norway, H2CO3 * K1 / Hplus)
norway$CO3 <- with(norway, K2 * HCO3 / Hplus)

norway$fCO2 <- with(norway, CO2 / K0)
norway$dfCO2 <- with(norway, fCO2 - fCO2.atm)

norway$OH <- with(norway, Kw/Hplus)

# Check
norway$DIC <- with(norway, H2CO3 + HCO3 + CO3)
norway$CA <- with(norway,  HCO3 + 2*CO3) 
```

```{r plot-carbonates}
ggplot(norway, aes(x = pH))+geom_point(aes(y = H2CO3, col = "H2CO3", size = DIC))+
  geom_point(aes(y = HCO3, col = "HCO3", size = DIC))+
  geom_point(aes(y = CO3, col = "CO3", size = DIC))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(y= "Concentration in mol/L", col = "Carbonate species", size = "DIC in mol/L")+
  theme_minimal()
```

```{r plot-fCO2}

ggplot(norway, aes(x = TOC, y = fCO2, col = survey))+
  geom_hline(yintercept = fCO2.atm.2019, col = "firebrick")+
  geom_hline(yintercept = fCO2.atm.2004, col = "dodgerblue3")+
  scale_color_manual(values = c("firebrick","dodgerblue3"))+
  #coord_trans(x = "log10")+
  geom_point()+
  labs(y = "pCO2, atm", x = "TOC, mol/L")+
  theme_minimal()

```

## Carbonate speciation from TA

The concentration of each carbonate species is then calculated based on total alkalinity as a proxy for $HCO_3^- + 2 \times CO_3^{2-}$. The resulting pCO2 is also calculated. 

```{r carbonate-speciation-from-TA}

norway$ta_CO3 <- with(norway,(TA - OH + Hplus) / (Hplus/K2 +2)) # in mol/L
norway$ta_HCO3 <- with(norway, ta_CO3*Hplus/K2)
norway$ta_H2CO3 <- with(norway, ta_HCO3*Hplus/K1)
norway$ta_fCO2 <- with(norway,ta_H2CO3/K0)

norway$ta_DIC <- with(norway, ta_H2CO3 + ta_HCO3 + ta_CO3)
```

## Save test dataset

The dataset is saved under the name "norway".

```{r save-norway}
saveRDS(norway,"norway.rds")
```

# Northern Lakes Survey

The dataset from the Northern Europeal Lakes Survey @Henriksen1998 is used as a training dataset. Additionnaly, it constitutes the basis for calculating national evasion rates in Norway, Sweden and Finland. 

## Download from NOFA database

The data is available in the NOFA database [@ninanor], including catchment polygons. 

Measurements methods for the Northern Lakes Survey are detailed in @Henriksen1996. Alkalinity is detailed in Annex 3. "alk_e" is a the variable corresponding to the alkalinity calibrated, taking into account the account the different methods used in Norway, Sweden and Finland. 


```{r load-data, eval = F}
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(),user = "camille.crapart", password = "camille",host = "vm-srv-wallace.vm.ntnu.no", dbname = "nofa")

ion <- tbl(con, sql("SELECT gid, ebint,nation,date,temp_c,latitude,longitude,ph,alk_e_ueq_l,alk_subst,k25_ms_m,ca_ueq_l,mg_ueq_l,na_ueq_l,k_ueq_l,nh4_ueq_l,cl_ueq_l,so4_ueq_l,no3_ueq_l,f_ueq_l,tot_p_ug_l,tot_n_ug_l,toc_mg_l,dist_closest_ebint,dist_2nd_closest_ebint FROM environmental.north_euro_lake_surv_1995 ")) %>% as.data.frame()
names(ion) <- c("gid", "ebint","nation","date","temp_c","latitude","longitude","ph","alk_ueq","alk","cond","ca","mg","na","k","nh4","cl","so4","no3","f","tp","tn","toc","dist_closest_ebint","dist_2nd_closest_ebint")
saveRDS(ion,"ion.rds")

lakes <- st_read(con,query = "SELECT gid, ebint, geom, area, perimtot FROM environmental.ecco_biwa_lakes_v_0_1 WHERE ebint IN (SELECT ebint FROM environmental.north_euro_lake_surv_1995)")
saveRDS(lakes,"lakes.rds")

catchments <- st_read(con,query = "SELECT ebint, geom FROM catchments.lake_catchments WHERE ebint IN (SELECT ebint FROM environmental.north_euro_lake_surv_1995)") # in UTM33
saveRDS(catchments,"catchments.rds")
catchment.poly <- st_transform(catchments, crs = st_crs("EPSG:4326")) # converts in World Geodetic System 1984
saveRDS(catchment.poly,"catchment.poly.rds")
catchment.poly.corine <- st_transform(catchment.poly, crs = st_crs("EPSG:3035")) # converts to EU extended referential ETRS89-extended
saveRDS(catchment.poly.corine, "catchment.poly.corine.rds")
```
## Near surface wind speed

Wind speed data is acquired from @C3S2020. Lakes from the NLS survey are larger than for the CBA and N112 survey. Therefore, we averaged the wind speed for the year and for the total area of the lakes.

```{r nws, eval = F}
lakes <- readRDS("lakes.rds")

nws.files <- list.files(path = "NWS", pattern = "Wind_WFDE5_CRU_1995", full.names = T) 
nws.list.stacks <- lapply(nws.files, raster::stack) %>% lapply(raster::crop, c(0,35,55,73))
nws.list.means <- lapply(nws.list.stacks, raster::mean, na.rm = T)
saveRDS(nws.list.means, "nlss.nws.list.mean.rds")

nws.list.mean <- readRDS("nlss.nws.list.mean.rds")
nws.1995 <- nws.list.mean %>% raster::stack() %>% raster::mean()

annual.nws.1995 <- raster::extract(nws.1995, lakes,fun = mean, na.rm = T, df = T)
annual.nws.1995$ebint <- lakes$ebint

saveRDS(annual.nws.1995, "annual.nws.nls.rds")
```


## Merge data for NLS

We merged the different datasets related to the NLS. Lakes with TOC and TP concentration inferior to 0 were removed (we assumed measurement errors in these lakes).

```{r merge-df, eval = F}

ion <- readRDS("ion.rds")
lakes <- readRDS("lakes.rds")
nws <- readRDS("annual.nws.nls.rds") %>% setNames(c("ID","nws_m.s","gid","ebint"))

nls <- merge(ion, lakes, by = "ebint") %>% merge(nws, by= "ebint")

nls$uncertain <- ifelse(nls$dist_closest_ebint/nls$dist_2nd_closest_ebint > 0.5, FALSE, TRUE)
nls$uncertain[which(is.na(nls$uncertain)==TRUE)] <- FALSE

nls$temp_c[which(is.na(nls$temp_c)==T)] <- 4

nls <- nls %>% filter(toc >= 0) %>% filter(tp > 0) %>% filter(uncertain == FALSE) 

saveRDS(nls,"nls.rds")
```

A simplified dataframe is created, with a selection of variables matching the variables used in the CBA and N112 survey. TOC and alkalinity are converted in mol/L.

```{r selected-df}
nls <- readRDS("nls.rds")

nlss <- data.frame(lake.id = nls$ebint)

nlss$pH <- nls$ph
nlss$Hplus <- 10^(-nlss$pH)
nlss$TA <- nls$alk_ueq * 1e-6
nlss$TOC <- nls$toc / 12.011 * 1e-3
nlss$nws_m.s <- nls$nws_m.s

nlss$long <- nls$longitude
nlss$lat <- nls$latitude
nlss$fCO2.atm <- fCO2.atm.1995

nlss$temp_c <- nls$temp_c
nlss$temp_k <- nls$temp_c + 273.15

nlss$lake.area <- nls$area * 1e-6 # m2 to km2

nlss$nation <- nls$nation


```

The dissociation constants are calculated as previously. 

```{r dissociation-constants, eval = F}
nlss$Kw = with(nlss, exp(148.9802 - 13847.26/temp_k - 23.6521*log(temp_k))) # Dickson and Riley, 1979

#nlss$K0 <- with(nlss, 10^-((-2622.38/temp_k) + 15.5873 - 0.0178471*temp_k)) #Harned and Davis, 1943
nlss$K0 <- with(nlss, exp(-58.0931+90.5069*100/temp_k+22.2940*log(temp_k/100))) # Weiss 1974
nlss$K1 <- with(nlss, 10^-((3404.71/temp_k)- 14.8435 + 0.032786*temp_k)) # Harned and Davis, 1943
nlss$K2 <- with(nlss, 10^-((2902.39/temp_k) - 6.4980 + 0.02379*temp_k)) # Harned and Scholes, 1941


nlss$OH <- with(nlss, Kw/Hplus)
```

## Save NLS

The final dataframe is saved under the name "nlss".

```{r remove-duplicate}
duplicated.id <- nlss$lake.id[which(duplicated(nlss$lake.id) == T)]
nlss[nlss$lake.id %in% duplicated.id, ]
unique.nlss <- nlss %>% group_by(lake.id, nation) %>% summarise_all(median)
```

```{r save-nlss, eval = F}
saveRDS(unique.nlss, "nlss.rds")

```



