---
title: "pCO2"
format: html
editor: visual
---

```{r dataset}
library(AquaEnv)
library(dplyr)
library(ggplot2)

cba.sel <- readRDS("cba.sel.rds")
boxplot(cba.sel$TOC)
cba <- cba.sel %>% subset(TOC < 40)

```

Problems with dataset:

-   In the CBA dataset, alkalinity is supposed to be in meq/L (and I double-checked) but I have a doubt about it, since I only get the right order of magnitude in the results if I consider it to be in ueq/L.

-   After talking to Nico, i confirm that CO2 is the sum of all carbonates (so SumCO2 in aquaenv)

Reference for Kh, K1, K2: Jan Åberg & B. Wallin, 2014

Reference for Organic Alkalinity correction: Liu et al 2020

Carbonate equilibrium: Appelo et Postma 2005

```{r variables}

# Equilibrium constants
Temp <- cba$temp # in K
A1 <- -58.0931
A2 <- 90.5069
A3 <- 22.2940
Kh <- exp(A1+A2*100/Temp+A3*log(Temp/100))

K1 <- 10^((-3404.71/Temp)+14.844-0.033*Temp)
K2 <- 10^((-2902.39/Temp)+6.498-0.0238*Temp)

# CO2 partial pressure in atmosphere
fCO2.2019 <- 410.5 # in ppm
fCO2.2019.atm <- fCO2.2019 * 1e-6 # in atm

# Total inorganic carbon
DIC <- cba$CO2 * 1e-6 # from umol/L to mol/L

# Total organic carbon
TOC <- cba$TOC/12.011 * 1e-3 # mg/L to mol/L

# Temperature, pH, Alkalinity
temp <- cba$T
TA <- cba$alk * 1e-6 # convert from ueq/L to mol/L
pH <- cba$pH_Kje
Hplus <- 10^(-pH) 

# Organic alkalinity
OA_Liu <- cba$toc * 1e3 / 12.011 * 0.08 * 1e-6 # toc from mg/L to umol/L multiplied by fraction of acids -> gives OrgAlk in ueq/L --> converted in mol/L

```

```{r distribution}
hist(DIC)
hist(TOC)
hist(OA_Liu)
hist(TA)

plot(TA, OA_Liu)
abline(a=0, b = 1)

```

# Using aquaenv

## aquaenv without correction

```{r aquaenv-no-correction}

ae1 <- aquaenv(S=0, t=temp,pH = pH, TA=TA, SumCO2 = NULL, fCO2atm = fCO2.2019.atm)

plot(DIC, ae1$SumCO2)
abline(a=0, b=1)

plot(TOC, ae1$fCO2*1e6)
abline(a = fCO2.2019 , b = 0)

plot(ae1$fCO2*1e-6, ae1$CO2 * Kh)

```

Aquaenv overstimate DIC concentration

## aquaenv with correction

```{r aquaenv-no-correction}

ae2 <- aquaenv(S=0, t=temp,pH = pH, TA=(TA-OA_Liu), SumCO2 = NULL, fCO2atm = fCO2.2019.atm)

plot(DIC, ae1$SumCO2)
abline(a=0, b=1)

plot(TOC, ae1$fCO2*1e6)
abline(a = fCO2.2019 , b = 0)
```

No change

# Compute fCO2 from CO2

$TIC = [H_2CO_3] + [HCO_3^-]+[CO_3^{2-}]$

$[HCO_3^-] = K_1 \times [H_2CO_3] / [H^+]$

$[CO_3^{2-}] = K_2 \times [HCO_3^-] /[H^+] = K_2 / [H^+] \times K_1 \times [H_2CO_3] / [H^+]$

$TIC = [H_2CO_3] + K_1 \times [H_2CO_3]/[H^+] + K_2 /[H^+] \times K_1 \times [H_2CO_3] / [H^+]$

$TIC = [H_2CO_3] \times (1 + K_1 / [H^+] + K_1 \times K_2 / [H^+]^2 )$

$[H_2CO_3] = TIC \times 1/(1 + K_1/[H^+] + K_1 \times K_2 / [H^+]^2)$

Could also take Ca in consideration: https://link.springer.com/content/pdf/10.1007/s002160050832.pdf

```{r Manually-find-pCO2}
#  DIC = H2CO3 + HCO3 +CO3 
H2CO3 <- DIC*1/(1+(K1/Hplus)+(K1*K2/Hplus^2)) # in mol/L
HCO3 <- H2CO3 * K1 / Hplus
CO3 <- K2 * HCO3 / Hplus

# Check
C <- H2CO3 + HCO3 + CO3
plot(DIC,C)

# Check with pH

qplot()+geom_point(aes(pH, H2CO3, col = "H2CO3"))+
  geom_point(aes(x = pH, y = HCO3, col = "HCO3"))+
  geom_point(aes(x=pH, CO3, col = "CO3"))+
  scale_color_manual(values = c("firebrick","lightblue","orange"))+
  theme_minimal()

fCO2 <- H2CO3/Kh
plot(TOC, fCO2*1e6)
abline(a = fCO2.2019, b = 0)
```

# Find OrgAlk

```{r OA}
Ca <- cba$ca / 2 * 1e-6 #ueq/L to mol/L
A <- HCO3+2*CO3+2*Ca
plot(TA, A)
abline(a = 0, b = 1)

plot(OA_Liu, A)
```
