---
title: "pCO2"
format: html
editor: visual
---

```{r dataset}
library(AquaEnv)
library(dplyr)
library(ggplot2)

cba.sel <- readRDS("cba.sel.rds")
boxplot(cba.sel$TOC)
cba <- cba.sel %>% subset(TOC < 40)

```

Problems with dataset:

-   In the CBA dataset, alkalinity is supposed to be in meq/L (and I double-checked) but I have a doubt about it, since I only get the right order of magnitude in the results if I consider it to be in ueq/L.

-   After talking to Nico, i confirm that CO2 is the sum of all carbonates (so SumCO2 in aquaenv)

Reference for headspace method: Jan Åberg & B. Wallin, 2014

Reference for K0, K1, K2 and Kw: Harned et Davis, 1943

Reference for Organic Alkalinity correction: Liu et al 2020

Carbonate equilibrium: Appelo et Postma 2005

Temperature: here we use in-situ temperature but since pH and Alk were measured in the lab, we should maybe use storage temperature (4 degrees)

```{r variables}

# Equilibrium constants
Temp <- cba$temp # in K

Kw = exp(148.9802 - 13847.26/Temp - 23.6521*log(Temp)) # Dickson and Riley, 1979

K0 <- 10^-((-2622.38/Temp) + 15.5873 - 0.0178471*Temp) #Harned and Davis, 1943
K1 <- 10^-((3404.71/Temp)- 14.8435 + 0.032786*Temp) # Harned and Davis, 1943
K2 <- 10^-((2902.39/Temp) - 6.4980 + 0.02379*Temp) # Harned and Scholes, 1941

# CO2 partial pressure in atmosphere
fCO2.2019 <- 410.5 # in ppm
fCO2.2019.atm <- fCO2.2019 * 1e-6 # in atm

# Total inorganic carbon
DIC <- cba$CO2 * 1e-6 # from umol/L to mol/L

# Total organic carbon
TOC <- cba$TOC/12.011 * 1e-3 # mg/L to mol/L

# Temperature, pH, Alkalinity
temp <- cba$T
TA <- cba$alk * 1e-6 # convert from ueq/L to mol/L
pH <- cba$pH_Kje
Hplus <- 10^(-pH)

# Conductivity
EC <- cba$EC_Kje # uS/cm

# Organic alkalinity
OA_Liu <- cba$toc * 1e3 / 12.011 * 0.08 * 1e-6 # toc from mg/L to umol/L multiplied by fraction of acids -> gives OrgAlk in ueq/L --> converted in mol/L

```

```{r distribution}
hist(DIC)
hist(TOC)
hist(OA_Liu)
hist(TA)

plot(TA, OA_Liu)
abline(a=0, b = 1)

```

# Using aquaenv

## aquaenv without correction

```{r aquaenv-no-correction}

ae1 <- aquaenv(S=0, t=temp,pH = pH, TA=TA, SumCO2 = NULL, fCO2atm = fCO2.2019.atm)

plot(DIC, ae1$SumCO2)
abline(a=0, b=1)

plot(TOC, ae1$fCO2)
abline(a = fCO2.2019.atm , b = 0)

plot(ae1$fCO2, ae1$CO2 / K0)
abline(a = 0, b = 1)
```

Aquaenv overstimate DIC concentration

## aquaenv with correction

```{r aquaenv-Liu-correction}

ae2 <- aquaenv(S=0, t=temp,pH = pH, TA=(TA-OA_Liu), SumCO2 = NULL, fCO2atm = fCO2.2019.atm)

plot(DIC, ae1$SumCO2)
abline(a=0, b=1)

plot(TOC, ae1$fCO2)
abline(a = fCO2.2019.atm , b = 0)
```

No change

## Aquaenv using pH and CO2, to find TA

```{r aquaenv-TA}

ae3 <- aquaenv(S=0, t=temp,pH = pH, SumCO2 = DIC, fCO2atm = fCO2.2019.atm)

plot(TA, ae3$TA)
abline(a=0, b=1)
```

# Compute fCO2 from CO2 manually

## Theory

$TIC = [H_2CO_3] + [HCO_3^-]+[CO_3^{2-}]$

$[HCO_3^-] = K_1 \times [H_2CO_3] / [H^+]$

$[CO_3^{2-}] = K_2 \times [HCO_3^-] /[H^+] = K_2 / [H^+] \times K_1 \times [H_2CO_3] / [H^+]$

$TIC = [H_2CO_3] + K_1 \times [H_2CO_3]/[H^+] + K_2 /[H^+] \times K_1 \times [H_2CO_3] / [H^+]$

$TIC = [H_2CO_3] \times (1 + K_1 / [H^+] + K_1 \times K_2 / [H^+]^2 )$

$[H_2CO_3] = TIC \times 1/(1 + K_1/[H^+] + K_1 \times K_2 / [H^+]^2)$

And $[H_2CO_3] = pCO_2 \times K_0$, or $pCO_2 = [H_2CO_3] / K_0$

Could also take Ca in consideration, but probably not significant here: https://link.springer.com/content/pdf/10.1007/s002160050832.pdf

## Carbonate speciation

```{r manually-find-pCO2}
#  DIC = H2CO3 + HCO3 +CO3 
H2CO3 <- DIC*1/(1+(K1/Hplus)+(K1*K2/Hplus^2)) # in mol/L
HCO3 <- H2CO3 * K1 / Hplus
CO3 <- K2 * HCO3 / Hplus

# Check
C <- H2CO3 + HCO3 + CO3
plot(DIC,C)

# Check with pH

qplot()+geom_point(aes(pH, H2CO3, col = "H2CO3"))+
  geom_point(aes(x = pH, y = HCO3, col = "HCO3"))+
  geom_point(aes(x=pH, CO3, col = "CO3"))+
  scale_color_manual(values = c("firebrick","lightblue","orange"))+
  theme_minimal()

fCO2 <- H2CO3/K0
plot(TOC, fCO2)
abline(a = fCO2.2019.atm, b = 0)
```

Less lakes are supersaturated compared to the aquaenv model. Less high fCO2 values

## Find Carbonate and Organic Alkalinity

$TA = [HCO_3^-] + 2[CO_3^{2-}] + [OH^-] - [H^+] + [other weak anions]$

Here, "other weak anions" come from organic matter // organic alkalnity.

```{r OA}
OH <- Kw/Hplus

#carbonate alkalinity
CA <- HCO3+2*CO3+OH-Hplus
plot(TA, CA)

abline(a = 0, b = 1)

#organic alkalnity
OA <- TA - CA
plot(TA, OA)
abline(a = 0, b = 1)
```

Organic alkalinity contributes a lot to total alkalinity.

From @Kerr2021: \* There are more errors in carbonate alkalinity in low salinity waters because the equilibrium constants have not been rigorously evaluated @Woosley2021 \* Alkalinity on filtered water \<- no microbial alkalinity. Contribution expected to be low @Kim2996, @Ko2006

## Compare aquaenv with manual

```{r compare-CA-aquaenv-manual}
qplot()+geom_point(aes(x=TA, y = ae3$TA, col = "AquaEnv"))+geom_point(aes(x = TA, y = CA, col = "Manual"))+
  scale_color_manual(values = c("firebrick","lightblue"))+labs(x = "TA, mol/L", y = "CA, mol/L")+
  theme_minimal()
```

Good correspondance between CA calculated with aquaenv and CA calculated manually!

# Model OA

```{r robust-model-OA}
library(MASS)
summary(m <- rlm(OA~TOC, method = "MM")) 
plot(TOC, OA)
abline(a = coef(m)[1], b = coef(m)[2], col = "red")
```

```{r lm-OA}
summary(m2 <- lm(OA~TOC))
plot(TOC, OA)
abline(a = coef(m2)[1], b = coef(m2)[2], col = "red")

summary(m2b <- lm(OA~EC))
plot(EC,OA)
abline(a = coef(m2b)[1], b = coef(m2b)[2], col = "red")

summary(m2c <- lm(OA~TOC+EC))
summary(m2d <- lm(OA~TOC*EC))
summary(m2e <- lm(OA~TOC:EC))
summary(m2f <- lm(OA~TOC*EC+TA))

OA_pred <- predict(m2f, data.frame(EC=EC, TOC = TOC, TA = TA), type = "response")
plot(OA,OA_pred)
abline(a = 0, b = 1, col = "red")
plot(TA,OA_pred)
abline(a = 0, b = 1, col = "red")
```

```{r test-pred-OA}
ae4 <- aquaenv(pH = pH, TA = (TA-OA_pred), t = temp, S = 0, SumCO2 = NULL)
plot(fCO2, ae4$fCO2)

```

```{r lm-OA-TA}
summary(m3 <- lm(OA~TA))
plot(TA,OA)
abline(a = coef(m3)[1], b = coef(m3)[2], col = "red")

OA_p <- predict(m3, data.frame(TA = TA))
plot(OA,OA_p)
plot(TA,OA_p)
abline(a = 0, b = 1, col = "red")
```

Predicts that OA is \~85% of TA

```{r aquaenv-OA_p}
ae5 <- aquaenv(pH = pH, TA = (TA-OA_p), t = temp, S=0, SumCO2 = NULL)

plot(fCO2, ae5$fCO2)
abline(a = 0, b = 1, col = "red")
```

# model logOA

from @Kulinski2012

$OA = \frac{K_{DOM} \times f.DOC}{[H^+]+K_{DOM}}$

OA depends on a dissociation constant $K_{DOM}$, pH, and the fraction of DOM that carries carboxylic group.

$log(OA) = log(K_{DOM})+log(f)+log(DOC)-log([H^+]+K_{DOM})$

From linear model:

$log(OA) = 0.63 \times log(TOC) + 1.06 \times pH - 11.76$

Then $log([H^+]-K_{DOM}) = - log([H^+]) \times 1,06 = -log([H^+] + exp(1.06))$

So $K_{DOM} = exp(1.06) =  2.88$

Constant terms are $log(K_{DOM}) + log(f)$ and should be equal to the intercept.

$log(2.88)+log(f) = -11.76$

$log(f) = -11.76 - 1.06$

$log(f) = -12.82$

$f = 2,7 \times 10^{-6}$

```{r log-OA}
logOA <- log(OA)
logTOC <- log(TOC)

summary(llm <- lm(logOA~logTOC+pH))

lOA <- 0.63*logTOC+1.06*pH-11.76

qplot(logOA,lOA)+geom_abline(intercept = 0, slope = 1)+
  theme_minimal()
``
```

According to @Kulinski2012:

-   K = 2.98e-8

-   f = 0.14

Tests are made with different values for K and f

```{r test-K-f}
Kk <- 2.98e-8
fk <- 0.14

OAk <- Kk*fk*TOC/(Hplus + Kk)
plot(OA, OAk)
cor(OA,OAk)

Ks <- seq(from = 1e-5, to = 3e-5, by = 0.5e-5)

fs <- seq(from = 0.1, to = 0.2, by = 0.01)

nb_test <- length(Ks)*length(fs)


OA_test <- data.frame(OA = OA)
for(i in 1:length(Ks)){
  for (j in 1:length(fs)){
    oa <- Ks[i]*fs[j]*TOC/(Hplus + Ks[i])
    col = i*j+1
    oaname <- paste("k",i,"f",j, sep = "")
   
   OA_test[,oaname] <- oa
   #colnames(OA_test)[col] <- oaname
  }
}

plot(OA,TOC)

ggplot(OA_test,aes(x=OA))+geom_point(aes(y= k1f1, col = "k1f1"))+
  geom_point(aes(y= k2f1, col = "k2f1"))+
  geom_point(aes(y= k3f1, col = "k3f1"))+
  geom_point(aes(y= k4f1, col = "k4f1"))+
  geom_point(aes(y= k5f1, col = "k5f1"))+
  geom_abline(slope = 1, intercept = 0)+
  scale_color_manual(values = c("firebrick","lightblue","black","orange","darkolivegreen"))+
  theme_minimal()

ggplot(OA_test,aes(x=OA))+geom_point(aes(y= k1f2, col = "k1f2"))+
  geom_point(aes(y= k2f2, col = "k2f2"))+
  geom_point(aes(y= k3f2, col = "k3f2"))+
  geom_point(aes(y= k4f2, col = "k4f2"))+
  geom_point(aes(y= k5f2, col = "k5f2"))+
  geom_abline(slope = 1, intercept = 0)+
  scale_color_manual(values = c("firebrick","lightblue","black","orange","darkolivegreen"))+
  theme_minimal()

ggplot(OA_test,aes(x=OA))+geom_point(aes(y= k1f3, col = "k1f3"))+
  geom_point(aes(y= k2f3, col = "k2f3"))+
  geom_point(aes(y= k3f3, col = "k3f3"))+
  geom_point(aes(y= k4f3, col = "k4f3"))+
  geom_point(aes(y= k5f3, col = "k5f3"))+
  geom_abline(slope = 1, intercept = 0)+
  scale_color_manual(values = c("firebrick","lightblue","black","orange","darkolivegreen"))+
  theme_minimal()

ggplot(OA_test,aes(x=OA))+geom_point(aes(y= k1f4, col = "k1f4"))+
  geom_point(aes(y= k2f4, col = "k2f4"))+
  geom_point(aes(y= k3f4, col = "k3f4"))+
  geom_point(aes(y= k4f4, col = "k4f4"))+
  geom_point(aes(y= k5f4, col = "k5f4"))+
  geom_abline(slope = 1, intercept = 0)+
  scale_color_manual(values = c("firebrick","lightblue","black","orange","darkolivegreen"))+
  theme_minimal()

```

# Model CA

```{r lm-CA}
summary(m4 <- rlm(CA~TA, method = "MM"))
plot(TA, CA)
abline(a=coef(m4)[1], b = coef(m4)[2])

CA_pred <- predict(m4, cba, type = "response")
plot(CA, CA_pred)

summary(m5 <- rlm(CA~TOC, method = "MM"))
plot(TOC, CA)
abline(a=coef(m5)[1], b = coef(m5)[2])
```

# Model DIC/TA

```{r ratio}
RA <- DIC/TA
RCA <-  CA/TA
ROA <- OA / TA

plot(TOC, ROA)
```
