---
title: "Reviews"
author: "Camille Crapart"
date: "2025-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F, results = T, cache = F)
options(knitr.kable.NA = '')
```

```{r libraries}
library(ggplot2)
library(ggforce)
library(ggpubr)
library(cowplot)
library(gridExtra)
library(colorspace)

library(gamlss)

library(kableExtra)

library(rsq)

library(sp)
library(sf)
library(dplyr)

library(raster)

pCO2.atm.2020 <- 411.51 * 1e-6 # oct 2020
pCO2.atm.2019 <- 408.75 * 1e-6 # oct 2019
pCO2.atm.2004 <- 374.63 * 1e-6 # oct 2004
pCO2.atm.1995 <- 360.17 * 1e-6
```

```{r fig-num-function}
library(captioner)
fig_nums <- captioner() 
tab_nums <- captioner(prefix = "Table")
```


# CBA CO2 fluxes

## CBA dataset

```{r import-data}
cba_no_area <- readxl::read_xlsx("CBA_100Lakes_Master.xlsx")
bdg.niva <- readRDS("bdg.niva.rds") %>% dplyr::select(c(CBA_Lake_ID, lake_area_km2, basin_area_km2))
cba <- merge(cba_no_area,bdg.niva, by.x = "Lake_ID", by.y = "CBA_Lake_ID") %>%
  dplyr::select(c(Lake_ID, Long, Lat, Altitude, CBA_date, T, TOC, DOC, pH_Kje, Alkalinity, CO2, lake_area_km2, basin_area_km2)) %>%
  rename(TA = Alkalinity, temp_c = T, pH = pH_Kje, CO2_umol = CO2, date = CBA_date)


# units
cba$TOC_mol.L <- cba$TOC/12.011*1e-3 #from mg/L to mol/L
cba$TA_mol.L <- cba$TA*1e-3 # convert from meq/L to mol/L
cba$temp_K <- cba$temp_c + 273.15
cba$pCO2_atm  <- 408.75 * 1e-6 # oct 2019
cba$Hplus <- 10^(-cba$pH)
cba$H2CO3 <- as.numeric(cba$CO2)*1e-6 #from umol/L to mol/L

# Constants

cba$Kw <- with(cba, exp(148.9802 - 13847.26/temp_K - 23.6521*log(temp_K))) # Dickson and Riley, 1979

cba$K0 <- with(cba, exp(-58.0931+90.5069*100/temp_K+22.2940*log(temp_K/100))) # Weiss 1974
cba$K1 <- with(cba, 10^-((3404.71/temp_K)- 14.8435 + 0.032786*temp_K)) # Harned and Davis, 1943
cba$K2 <- with(cba, 10^-((2902.39/temp_K) - 6.4980 + 0.02379*temp_K)) # Harned and Scholes, 1941

# DIC
cba$HCO3 <- with(cba, H2CO3 * K1 / Hplus)
cba$CO3 <- with(cba, K2 * HCO3 / Hplus)
cba$DIC <- with(cba, H2CO3 + HCO3 + CO3)

#pCO2
cba$pCO2 <- with(cba, H2CO3 / K0)

#cba.spdf <- SpatialPointsDataFrame(coords = cba[,c("Long","Lat")], data = cba[,c("Long","Lat","Lake_ID")], proj4string = CRS("+proj=longlat +datum=WGS84"))

cba.sf <- st_as_sf(cba, coords = c("Long","Lat"))

st_crs(cba.sf) <- 4326
```

## Wind speed 
```{r wind-speed-nws, eval = F}
nws.files <- list.files(path = "NWS", pattern = "Wind_WFDE5_CRU_2019", full.names = T) 
nws.raster.crop <- lapply(nws.files, raster) %>% lapply(raster::crop, c(0,35,55,73)) # test with head

#nws.list.means <- lapply(nws.list.stacks, raster::mean, na.rm = T)
nws.mean <- nws.raster.crop %>% raster::stack() %>% raster::mean(na.rm = T)
nws.xiu <- lapply(nws.raster.crop, FUN = function(r1){(r1 - nws.mean)^2})

raster.sum <- function(x) {sum(x)}
nws.sum <- nws.xiu %>% raster::stack() %>% raster.sum()

sd.function <- function(x) {sqrt(sum(x)/length(nws.raster.crop))}
nws.sd <- nws.xiu %>% raster::stack() %>% sd.function()

cba.spdf <- SpatialPointsDataFrame(coords = cba[,c("Long","Lat")], data = cba[,c("Long","Lat","Lake_ID")], proj4string = CRS("+proj=longlat +datum=WGS84"))
nws.mean.cba <- raster::extract(nws.mean, cba.spdf, sp = T)
nws.mean.cba.df <- nws.mean.cba@data %>% setNames(c("Long","Lat","Lake_ID","nws_mean"))

nws.sd.cba <- raster::extract(nws.sd, cba.spdf, sp = T)
nws.sd.cba.df <- nws.sd.cba@data %>% setNames(c("Long","Lat","Lake_ID","nws_sd"))

cba.sf <- cba %>% merge(nws.mean.cba.df, by = intersect(names(cba), names(nws.mean.cba.df)))%>%
  merge(nws.sd.cba.df, by = intersect(names(cba), names(nws.sd.cba.df)))

```

```{r wind-speed-10-years-mean-cba, eval = T}
sd.function <- function(x) {sqrt(sum(x)/120)}

u10.2019 <- raster::stack("ERA5/u10_2014_2023.nc")
u10.2019.mean <- u10.2019 %>% mean(na.rm = T)
u10.2019.sd <- (u10.2019 - u10.2019.mean)^2 %>% raster::stack() %>% sd.function()

v10.2019 <- raster::stack("ERA5/v10_2014_2023.nc")
v10.2019.mean <- v10.2019 %>% mean(na.rm = T)
v10.2019.sd <- (v10.2019 - v10.2019.mean)^2 %>% raster::stack() %>% sd.function()

wind.2019.mean <- sqrt(u10.2019^2 + v10.2019^2) %>% mean()
wind.2019.sd <- (sqrt(u10.2019^2 + v10.2019^2) - wind.2019.mean)^2 %>% raster::stack() %>% sd.function()
 
wind.2019.cba <- raster::extract(wind.2019.mean, cba.sf)
cba.sf$wind_mean <- wind.2019.cba # pb with df: lake_ID disappears
#wind.2019.cba.df <- wind.2019.cba@data %>% setNames(c("Long","Lat","Lake_ID","wind_mean"))
 
wind.2019.sd.cba <- raster::extract(wind.2019.sd, cba.sf)
cba.sf$wind_sd <- wind.2019.sd.cba
#wind.2019.sd.cba.df <- wind.2019.sd.cba@data %>% setNames(c("Long","Lat","Lake_ID","wind_sd"))

#cba.sf <- cba %>% merge(wind.2019.cba.df, by = intersect(names(cba), names(wind.2019.cba.df)))%>%
# merge(wind.2019.sd.cba.df, by = intersect(names(cba), names(wind.2019.cba.df)))
```


## Compute k600 with monte-carlo
```{r k600-monte-carlo-cba}
cba.sf$Sc <- with(cba.sf, 1923.6-125.06*temp_c+4.3733*temp_c^2-0.086781*temp_c^3)

cba.sf$k600vp <- with(cba.sf, 2.51+1.48*wind_mean+0.39*wind_mean*log10(lake_area_km2)) # lake.area already in km2, in cm/h. Vachon and Prairie 2013
cba.sf$k600cc <- with(cba.sf, 2.07+0.25*wind_mean^1.7) # lake area in km2. Cole & Caraco 1998
cba.sf$k600mci <- with(cba.sf, 2+1.5*wind_mean) # @MacIntyre2020


cba.sf$k600vp_mcmean <- NA
cba.sf$k600vp_mcsd <- NA
cba.sf$k600cc_mcmean <- NA
cba.sf$k600cc_mcsd <- NA
cba.sf$k600mci_mcmean <- NA
cba.sf$k600mci_mcsd <- NA


for(i in 1:dim(cba.sf)[1]){
  lake_area_i <- cba.sf$lake_area_km2[i]
  Sc_i <- cba.sf$Sc[i]
  pCO2_i <- cba.sf$pCO2[i]
  K0_i <- cba.sf$K0[i]

  
  windm_i <- cba.sf$wind_mean[i]
  windsd_i <- cba.sf$wind_sd[i]
  wind <- runif(100000, windm_i - windsd_i, windm_i + windsd_i) 
  
  
  # k600 from Vachon & Prairie
  
  k600vp <- 2.51+1.48*wind+0.39*wind*log10(lake_area_i)
  k600vp_mi <- mean(k600vp)
  k600vp_sdi <- sd(k600vp)
  
  cba.sf$k600vp_mcmean[i] <- k600vp_mi
  cba.sf$k600vp_mcsd[i] <- k600vp_sdi

  
  ECO2vp <- k600vp*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2vp_mi <- mean(ECO2vp)
  ECO2vp_sdi <- sd(ECO2vp)
  
  cba.sf$ECO2vp_mcmean[i] <- ECO2vp_mi
  cba.sf$ECO2vp_mcsd[i] <- ECO2vp_sdi
  
  # k600 from Cole and Caraco
  
  k600cc <- 2.07+0.25*wind^1.7
  k600cc_mi <- mean(k600cc)
  k600cc_sdi <- sd(k600cc)

  cba.sf$k600cc_mcmean[i] <- k600cc_mi
  cba.sf$k600cc_mcsd[i] <- k600cc_sdi
  
  ECO2cc <- k600cc*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2cc_mi <- mean(ECO2cc)
  ECO2cc_sdi <- sd(ECO2cc)
  
  cba.sf$ECO2cc_mcmean[i] <- ECO2cc_mi
  cba.sf$ECO2cc_mcsd[i] <- ECO2cc_sdi
  
  # k600 from MacIntyre
  
  k600mci <- 2+1.5*wind
  k600mci_mi <- mean(k600mci)
  k600mci_sdi <- sd(k600mci)
  
  cba.sf$k600mci_mcmean[i] <- k600mci_mi
  cba.sf$k600mci_mcsd[i] <- k600mci_sdi
  
  ECO2mci <- k600mci*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2mci_mi <- mean(ECO2mci)
  ECO2mci_sdi <- sd(ECO2mci)
  
  cba.sf$ECO2mci_mcmean[i] <- ECO2mci_mi
  cba.sf$ECO2mci_mcsd[i] <- ECO2mci_sdi
  
}


ggplot(cba.sf)+
  geom_point(aes(y = ECO2vp_mcmean, x = as.character(Lake_ID), col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, x = as.character(Lake_ID), col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, x = as.character(Lake_ID), col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, x = as.character(Lake_ID), col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, x = as.character(Lake_ID), col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, x = as.character(Lake_ID), col = "MacIntyre"))+
  theme_minimal()
  
ggplot(cba.sf, aes(x = TOC))+
  geom_point(aes(y = ECO2vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, col = "MacIntyre"))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "TOC, mg/L", y = "CO2 evasion in gC/m2/day", col = "")+
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()

ggplot(cba.sf, aes(x = lake_area_km2))+
  geom_point(aes(y = ECO2vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, col = "MacIntyre"))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "Lake area, km2", y = "CO2 evasion in gC/m2/day", col = "")+
  scale_x_log10()+
  theme_minimal()


ggplot(cba.sf, aes(x = ECO2cc_mcmean, y = ECO2mci_mcmean))+
  geom_point()+geom_abline(slope = 1, intercept = 0)

```


# N112

## N112 dataset

```{r import-data-n112}
n112.all <- read.table("N112_subset_110909.txt", sep = " ", header = TRUE, na.strings = c(".", NA))

n112 <- n112.all %>% subset(ykoord > 6e6) %>% subset(is.na(TIC.ug) == F) %>% subset(is.na(alkendpoint) == F) %>% 
  dplyr::select(c("xkoord", "ykoord","hoh.NVE","Vatnlnr","lake.area","poly.area", "TOC", "alkendpoint","temp.situ","pH","co2.umolar","co2.uatm")) %>%
  rename(Long = xkoord, Lat = ykoord, Lake_ID = Vatnlnr, lake_area_km2 = lake.area, basin_area_km2 = poly.area, TA = alkendpoint, temp_c = temp.situ, CO2_umol = co2.umolar, Altitude = hoh.NVE)

# units
n112$TOC_mol.L <- n112$TOC/12.011*1e-3 #from mg/L to mol/L
n112$TA_mol.L <- n112$TA*1e-6 # convert from ueq/L to mol/L
n112$temp_K <- n112$temp_c + 273.15
n112$pCO2_atm  <- 374.63 * 1e-6 # oct 2004
n112$Hplus <- 10^(-n112$pH)
n112$H2CO3 <- as.numeric(n112$CO2_umol)*1e-6 #from umol/L to mol/L

# Constants

n112$Kw <- with(n112, exp(148.9802 - 13847.26/temp_K - 23.6521*log(temp_K))) # Dickson and Riley, 1979

n112$K0 <- with(n112, exp(-58.0931+90.5069*100/temp_K+22.2940*log(temp_K/100))) # Weiss 1974
n112$K1 <- with(n112, 10^-((3404.71/temp_K)- 14.8435 + 0.032786*temp_K)) # Harned and Davis, 1943
n112$K2 <- with(n112, 10^-((2902.39/temp_K) - 6.4980 + 0.02379*temp_K)) # Harned and Scholes, 1941

# DIC
n112$HCO3 <- with(n112, H2CO3 * K1 / Hplus)
n112$CO3 <- with(n112, K2 * HCO3 / Hplus)
n112$DIC <- with(n112, H2CO3 + HCO3 + CO3)

#pCO2
n112$pCO2 <- with(n112, H2CO3 / K0)
# 
# n112.spdf <- SpatialPointsDataFrame(coords = n112[,c("Long","Lat")], data = n112[,c("Long","Lat","Lake_ID")], proj4string = CRS("+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
# 
# n112.spdf.wgs <- n112.spdf %>% spTransform(CRS("+proj=longlat +datum=WGS84"))


n112.sf <- n112 %>% st_as_sf(coords = c("Long","Lat")) 
st_crs(n112.sf) <- 25832
n112.sf <- n112.sf %>% st_transform(crs = 4326)
```

## Wind speed

```{r wind-speed-10-years-mean-n112, eval = T}
sd.function <- function(x) {sqrt(sum(x)/120)}

u10.2004 <- raster::stack("ERA5/u10_1999_2008.nc")
u10.2004.mean <- u10.2004 %>% mean(na.rm = T)
u10.2004.sd <- (u10.2004 - u10.2004.mean)^2 %>% stack %>% sd.function()

v10.2004 <- raster::stack("ERA5/v10_1999_2008.nc")
v10.2004.mean <- v10.2004 %>% mean(na.rm = T)
v10.2004.sd <- (v10.2004 - v10.2004.mean)^2 %>% stack %>% sd.function()

wind.2004.mean <- sqrt(u10.2004^2 + v10.2004^2) %>% mean()
wind.2004.sd <- (sqrt(u10.2004^2 + v10.2004^2) - wind.2004.mean)^2 %>% stack %>% sd.function()

wind.2004.n112 <- raster::extract(wind.2004.mean, n112.sf)
n112.sf$wind_mean <- wind.2004.n112 
#wind.2004.n112.df <- wind.2004.n112@data %>% setNames(c("Long","Lat","Lake_ID","wind_mean"))

wind.2004.sd.n112 <- raster::extract(wind.2004.sd, n112.sf)
n112.sf$wind_sd <- wind.2004.sd.n112
#wind.2004.sd.n112.df <- wind.2004.sd.n112@data %>% setNames(c("Long","Lat","Lake_ID","wind_sd"))

#n112.wind <- n112 %>% merge(wind.2004.n112.df, by = intersect(names(n112), names(wind.2004.n112.df)))%>%
#  merge(wind.2004.sd.n112.df, by = intersect(names(n112), names(wind.2004.n112.df)))
```

## Compute k600 with Monte Carlo

```{r k600-monte-carlo-n112}
n112.sf$Sc <- with(n112.sf, 1923.6-125.06*temp_c+4.3733*temp_c^2-0.086781*temp_c^3)

n112.sf$k600vp <- with(n112.sf, 2.51+1.48*wind_mean+0.39*wind_mean*log10(lake_area_km2)) # lake.area already in km2, in cm/h. Vachon and Prairie 2013
n112.sf$k600cc <- with(n112.sf, 2.07+0.25*wind_mean^1.7) # lake area in km2. Cole & Caraco 1998
n112.sf$k600mci <- with(n112.sf, 2+1.5*wind_mean) # @MacIntyre2020


n112.sf$k600vp_mcmean <- NA
n112.sf$k600vp_mcsd <- NA
n112.sf$k600cc_mcmean <- NA
n112.sf$k600cc_mcsd <- NA
n112.sf$k600mci_mcmean <- NA
n112.sf$k600mci_mcsd <- NA


for(i in 1:dim(n112.sf)[1]){
  lake_area_i <- n112.sf$lake_area_km2[i]
  Sc_i <- n112.sf$Sc[i]
  pCO2_i <- n112.sf$pCO2[i]
  K0_i <- n112.sf$K0[i]

  
  windm_i <- n112.sf$wind_mean[i]
  windsd_i <- n112.sf$wind_sd[i]
  wind <- runif(100000, windm_i - windsd_i, windm_i + windsd_i) 
  
  
  # k600 from Vachon & Prairie
  
  k600vp <- 2.51+1.48*wind+0.39*wind*log10(lake_area_i)
  k600vp_mi <- mean(k600vp)
  k600vp_sdi <- sd(k600vp)
  
  n112.sf$k600vp_mcmean[i] <- k600vp_mi
  n112.sf$k600vp_mcsd[i] <- k600vp_sdi

  
  ECO2vp <- k600vp*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2vp_mi <- mean(ECO2vp)
  ECO2vp_sdi <- sd(ECO2vp)
  
  n112.sf$ECO2vp_mcmean[i] <- ECO2vp_mi
  n112.sf$ECO2vp_mcsd[i] <- ECO2vp_sdi
  
  # k600 from Cole and Caraco
  
  k600cc <- 2.07+0.25*wind^1.7
  k600cc_mi <- mean(k600cc)
  k600cc_sdi <- sd(k600cc)

  n112.sf$k600cc_mcmean[i] <- k600cc_mi
  n112.sf$k600cc_mcsd[i] <- k600cc_sdi
  
  ECO2cc <- k600cc*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2cc_mi <- mean(ECO2cc)
  ECO2cc_sdi <- sd(ECO2cc)
  
  n112.sf$ECO2cc_mcmean[i] <- ECO2cc_mi
  n112.sf$ECO2cc_mcsd[i] <- ECO2cc_sdi
  
  # k600 from MacIntyre
  
  k600mci <- 2+1.5*wind
  k600mci_mi <- mean(k600mci)
  k600mci_sdi <- sd(k600mci)
  
  n112.sf$k600mci_mcmean[i] <- k600mci_mi
  n112.sf$k600mci_mcsd[i] <- k600mci_sdi
  
  ECO2mci <- k600mci*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2mci_mi <- mean(ECO2mci)
  ECO2mci_sdi <- sd(ECO2mci)
  
  n112.sf$ECO2mci_mcmean[i] <- ECO2mci_mi
  n112.sf$ECO2mci_mcsd[i] <- ECO2mci_sdi
  
}


ggplot(n112.sf)+
  geom_point(aes(y = ECO2vp_mcmean, x = as.character(Lake_ID), col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, x = as.character(Lake_ID), col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, x = as.character(Lake_ID), col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, x = as.character(Lake_ID), col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, x = as.character(Lake_ID), col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, x = as.character(Lake_ID), col = "MacIntyre"))+  
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "TOC, mg/L", y = "CO2 evasion in gC/m2/day", col = "")+
  theme_minimal()
  
ggplot(n112.sf, aes(x = TOC))+
  geom_point(aes(y = ECO2vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, col = "MacIntyre"))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "TOC, mg/L", y = "CO2 evasion in gC/m2/day", col = "")+
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()

ggplot(n112.sf, aes(x = lake_area_km2))+
  geom_point(aes(y = ECO2vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, col = "MacIntyre"))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "Lake area, km2", y = "CO2 evasion in gC/m2/day", col = "")+
  scale_x_log10()+
  theme_minimal()


ggplot(n112.sf, aes(x = ECO2cc_mcmean, y = ECO2mci_mcmean))+
  geom_point()+geom_abline(slope = 1, intercept = 0)

```

# Predict pCO2 from TOC

## Norway dataset


```{r remove-outliers-cba}
ggarrange(
ggplot(cba.sf) + geom_boxplot(aes(x = "TOC, g/L", y = TOC))+labs(y = ""),
ggplot(cba.sf) + geom_boxplot(aes(x = "pCO2, uatm", y = pCO2*1e6))+labs(y = ""),
ggplot(cba.sf) +geom_boxplot(aes(x = "pH", y = pH))+labs(y = ""),
ncol = 3
)

cba.w <- cba.sf %>% subset(TOC < 30 & pCO2 < 0.004)

ggarrange(
ggplot(cba.w) + geom_boxplot(aes(x = "TOC, g/L", y = TOC))+labs(y = ""),
ggplot(cba.w) + geom_boxplot(aes(x = "pCO2, uatm", y = pCO2*1e6))+labs(y = ""),
ggplot(cba.w) +geom_boxplot(aes(x = "pH", y = pH))+labs(y = ""),
ncol = 3
)

ggarrange(
ggplot(cba.w) + geom_density(aes(x = TOC))+labs(y = "", x = "TOC, mg/L"),
ggplot(cba.w) + geom_density(aes(x = pCO2*1e6))+labs(y = "", x = "pCO2, uatm"),
ggplot(cba.w) +geom_density(aes(x = pH))+labs(y = ""),
nrow = 3
)

```

```{r explore-relations}
ggplot(cba.w)+geom_point(aes(x=TOC_mol.L, y = pCO2)) + 
  geom_abline(slope = 1, intercept = pCO2.atm.2019)+
  labs(title = "CBA")

ggplot(n112.sf)+geom_point(aes(x=TOC_mol.L, y = pCO2))+ 
  geom_abline(slope = 1, intercept = pCO2.atm.2004)+
  labs(title = "n112")
  
```

```{r bind-cba-n112}
setdiff(names(cba.w), names(n112.sf))
setdiff(names(n112.sf), names(cba.w))

cba.tobind <- dplyr::select(cba.w, -c("date","DOC")) %>% mutate(survey = "CBA")
n112.tobind <- dplyr::select(n112.sf, -"co2.uatm") %>% mutate(survey = "N112")
n112.tobind <- n112.tobind[,names(cba.tobind)]

norway <- rbind(cba.tobind, n112.tobind)
```

## Test computing pCO2 using TA

```{r pCO2-from-ta}
norway$OH <- with(norway, Kw/Hplus)

norway$ta_CO3 <- with(norway,(TA_mol.L - OH + Hplus)/(Hplus/K2 +2)) # in molL
norway$ta_HCO3 <- with(norway, ta_CO3*Hplus/K2)
norway$ta_H2CO3 <- with(norway, ta_HCO3*Hplus/K1)
norway$ta_pCO2 <- with(norway,ta_H2CO3/K0)

norway$ta_DIC <- with(norway, ta_H2CO3 + ta_HCO3 + ta_CO3)
```

```{r pCO2-with-measured-ta}

ggplot(norway)+geom_point(aes(x = pCO2, y = ta_pCO2, col = survey), alpha = 0.5)+
  scale_color_manual(values = c( "firebrick","dodgerblue3"))+
  theme_bw(base_size = 15)+
  theme(legend.position = "bottom")

ggplot(norway,aes(x= pH))+geom_point(aes(y = ta_pCO2*1e6, col = "pCO2 from TA"), alpha = 0.5)+
  geom_point(aes(y = pCO2*1e6, col = "Measured pCO2"), alpha = 0.5)+
  labs(y = "pCO2, μatm", x = "pH", col = "")+
  scale_color_manual(values = c( "firebrick","dodgerblue3"))+
  facet_grid(cols = vars(survey))+
  theme_bw(base_size = 15)+
  theme(legend.position = "bottom")

with(norway, cor(ta_pCO2, pCO2, use = "pairwise.complete.obs")^2)
with(subset(norway, pH > 5.4), cor(ta_pCO2, pCO2, use = "pairwise.complete.obs")^2)

```

## Test computing pCO2 using TA, accounting for organic alkalinity

```{r hruska}
pKa1 <- 3.04
Ka1 <- 10^(-pKa1)

pKa2 <- 4.42
Ka2 <- 10^(-pKa2)

pKa3 <- 6.46
Ka3 <- 10^(-pKa3)

m <- 10.2*1e-6*12.011*1e3*1e-3# ueq/mg to eq/mol, site density BUT THERE IS A 1e3 PB: have to multiply by 1e-3 to stay in the range
norway$RCOO <- with(norway, (Ka1/Hplus + 2*Ka2/Hplus + 3*Ka3/Hplus)* m/3 * TOC) # estimated in eq/L

norway$RCOO_4.5 <- with(norway, (Ka1/10^(-4.5) + 2*Ka2/10^(-4.5) + 3*Ka3/10^(-4.5))* m/3 * TOC)

norway$OA_hruska <- with(norway, RCOO-RCOO_4.5)

ggplot(norway)+
  geom_point(aes(x = pH, y = RCOO, col = survey))+
  scale_color_manual(values = c("firebrick","dodgerblue3"))+
  labs(y="[RCOO-], mol/L")+
  theme_minimal()+
  theme(legend.position = "bottom")

norway$CA_hruska <- with(norway, TA - OA_hruska - OH + Hplus) 

norway$CO3_hruska <- with(norway,CA_hruska / (Hplus/K2 +2)) # in mol/L
norway$HCO3_hruska <- with(norway, CO3_hruska*Hplus/K2)
norway$H2CO3_hruska <- with(norway, HCO3_hruska*Hplus/K1)
norway$pCO2_hruska <- with(norway,H2CO3_hruska/K0)

with(norway, cor(pCO2_hruska, pCO2, use = "pairwise.complete.obs")^2)
with(subset(norway, pH > 5.4), cor(pCO2_hruska, pCO2, use = "pairwise.complete.obs")^2)

ggplot(norway,aes(x= pH))+
  geom_point(aes(y = pCO2_hruska*1e6, col = "pCO2 from calculated from Hruska et al"), alpha = 0.5)+
  geom_point(aes(y = pCO2*1e6, col = "Measured pCO2"), alpha = 0.5)+
  labs(y = "pCO2, uatm", x = "pH", col = "")+
  scale_color_manual(values = c( "firebrick","dodgerblue3"))+
  facet_grid(cols = vars(survey), scales = "free_y")+
  theme_minimal(base_size = 15)+
  theme(legend.position = "bottom")

ggplot(subset(norway, survey == "CBA"),aes(x= pH))+
  geom_point(aes(y = pCO2_hruska*1e6, col = "pCO2 from calculated from Hruska et al"), alpha = 0.5)+
  geom_point(aes(y = pCO2*1e6, col = "Measured pCO2"), alpha = 0.5)+
  labs(y = "pCO2, uatm", x = "pH", col = "")+
  scale_color_manual(values = c( "firebrick","dodgerblue3"))+
  facet_grid(cols = vars(survey), scales = "free_y")+
  theme_minimal(base_size = 15)+
  theme(legend.position = "bottom")
```

```{r liu}

norway$OA_Liu <- 0.08 * norway$TOC
norway$CA_Liu <- with(norway, TA-OA_Liu - OH + Hplus)

norway$CO3_Liu <- with(norway,CA_Liu / (Hplus/K2 +2)) # in mol/L
norway$HCO3_Liu <- with(norway, CO3_Liu*Hplus/K2)
norway$H2CO3_Liu <- with(norway, HCO3_Liu*Hplus/K1)
norway$pCO2_Liu <- with(norway,H2CO3_Liu/K0)

norway$DIC_Liu <- with(norway, H2CO3_Liu + HCO3_Liu + CO3_Liu)

ggplot(norway,aes(x= pH))+geom_point(aes(y = pCO2_Liu, col = "pCO2 from calculated from Liu et al."), alpha = 0.5)+
  geom_point(aes(y = pCO2*1e6, col = "Measured pCO2"), alpha = 0.5)+
  labs(y = "pCO2, atm", x = "pH", col = "")+
  scale_color_manual(values = c( "firebrick","dodgerblue3"))+
  facet_grid(cols = vars(survey))+
  theme_minimal(base_size = 15)+
  theme(legend.position = "bottom")


with(norway, cor(pCO2_Liu, pCO2, use = "pairwise.complete.obs")^2)
with(subset(norway, pH > 5.4), cor(pCO2_Liu, pCO2, use = "pairwise.complete.obs")^2)
```

## Model pCO2 ~ TOC 

### Train and test sets

```{r test-predict-pco2-norway-train-seed1}

set.seed(1)
norway.train <- norway %>% slice_sample(prop = 0.7)
norway.test <- norway %>% filter(!Lake_ID %in% norway.train$Lake_ID)


ggplot(norway.train)+geom_point(aes(x=TOC_mol.L, y = pCO2, col = survey)) + 
  geom_abline(slope = 1, intercept = pCO2.atm.2019)+
  labs(title = "norway.train")

ggplot(norway.test)+geom_point(aes(x=TOC_mol.L, y = pCO2, col = survey))+ 
  geom_abline(slope = 1, intercept = pCO2.atm.2004)+
  labs(title = "n112")

pco2.toc.train <- glm(formula = pCO2~TOC_mol.L+0+offset(pCO2_atm), data = norway.train, family = Gamma(link = "identity"))

pco2.toc.train.sum <- summary(pco2.toc.train)$coefficients %>% as.data.frame() 
pco2.toc.train.sp <- rownames(pco2.toc.train.sum)[which(pco2.toc.train.sum$`Pr(>|t|)` < 0.05)] %>% paste(collapse = ", ")
pco2.toc.train.sum <- pco2.toc.train.sum %>% format(scientific = T, digits = 2)

norway.train$pCO2_pred <- pco2.toc.train$fitted.values

kable(pco2.toc.train.sum, digits = 2) %>% kable_styling(bootstrap_options = "bordered")

norway.test$pCO2_pred <- predict(pco2.toc.train, newdata = norway.test, type = "response")

ggplot(norway.test) + geom_point(aes(x = pCO2*1e6, y = pCO2_pred*1e6, col = survey))+
    geom_text(x = 500, y = 2000,label = paste0('r2 = ', round((cor(norway.test$pCO2,norway.test$pCO2_pred))^2,2)))+
  labs(x = "pCO2 measured, atm", y = "pCO2 predicted, atm", title = "pCO2 predicted on test set, model fitted on train set")+
  geom_abline(slope = 1, intercept = 0, col = "darkblue")+
  theme_minimal()

```


### Cross-validation

```{r predict-pco2-on-full-dataset-with-cv-of-coefficient}
pco2.toc.nor <- glm(formula = pCO2~TOC_mol.L+0+offset(pCO2_atm), data = norway, family = Gamma(link = "identity"))

pco2.toc.nor.sum <- summary(pco2.toc.nor)$coefficients %>% as.data.frame() 
pco2.toc.train.sp <- rownames(pco2.toc.nor.sum)[which(pco2.toc.nor.sum$`Pr(>|t|)` < 0.05)] %>% paste(collapse = ", ")
pco2.toc.nor.sum <- pco2.toc.nor.sum %>% format(scientific = T, digits = 2)

norway$pCO2_pred <- pco2.toc.nor$fitted.values

kable(pco2.toc.nor.sum, digits = 2) %>% kable_styling(bootstrap_options = "bordered")

norway$pCO2_pred <- predict(pco2.toc.nor, newdata = norway, type = "response")
norway$pCO2_pred.se <- predict(pco2.toc.nor, newdata = norway, type = "response", se.fit = T)$se.fit

ggplot(norway) + geom_point(aes(x = pCO2*1e6, y = pCO2_pred*1e6, col = survey))+
    geom_text(x = 500, y = 2000,label = paste0('r2 = ', round((cor(norway$pCO2,norway$pCO2_pred))^2,2)))+
  labs(x = "pCO2 measured, atm", y = "pCO2 predicted, atm", title = "pCO2 predicted on test set, model fitted on train set")+
  geom_abline(slope = 1, intercept = 0, col = "darkblue")+
  theme_minimal()

```

```{r check-robustness-with-cv}

library("cv")

cv.nor <- cv(pco2.toc.nor) 
cv.nor.loo <- cv(pco2.toc.nor, k = "loo")

cv.nor.coefs <- cv.nor$details$coefficients %>% as.data.frame() %>% t() %>% as.data.frame() %>% rename(coef = TOC_mol.L) 
cv.nor.coefs$fold <- rownames(cv.nor.coefs) %>% grep(pattern = '[:digit:]')

plot(coef~fold, cv.nor.coefs)


print(mean(cv.nor.coefs$coef))

```
```{r pearson}
print(paste("r between pCO2 predicted with TA and pCO2 measured = ", round(cor(norway$pCO2, norway$ta_pCO2, use = "pairwise.complete"), 2)))
print(paste("r between pCO2 predicted with TA and pCO2 measured, pH > 5.4 = ", round(cor(subset(norway, pH > 5.4)$pCO2, subset(norway, pH > 5.4)$ta_pCO2, use = "pairwise.complete"), 2)))
print(paste("r between pCO2 predicted with TA and pCO2 measured, pH < 5.4 = ", round(cor(subset(norway, pH < 5.4)$pCO2, subset(norway, pH < 5.4)$ta_pCO2, use = "pairwise.complete"), 2)))
print(paste("r between pCO2 predicted with TOC and pCO2 measured = ", round(cor(norway$pCO2, norway$pCO2_pred, use = "pairwise.complete"), 2)))
print(paste("r between pCO2 predicted with TOC and pCO2 measured, pH > 5.4 = ", round(cor(subset(norway, pH > 5.4)$pCO2, subset(norway, pH > 5.4)$pCO2_pred, use = "pairwise.complete"), 2)))
print(paste("r between pCO2 predicted with TOC and pCO2 measured, pH < 5.4 = ", round(cor(subset(norway, pH < 5.4)$pCO2, subset(norway, pH < 5.4)$pCO2_pred, use = "pairwise.complete"), 2)))
```

# ECO2 for norway dataset

## Compute k600 with Monte-Carlo

```{r k600-monte-carlo-norway}
norway$Sc <- with(norway, 1923.6-125.06*temp_c+4.3733*temp_c^2-0.086781*temp_c^3)

norway$k600vp <- with(norway, 2.51+1.48*wind_mean+0.39*wind_mean*log10(lake_area_km2)) # lake.area already in km2, in cm/h. Vachon and Prairie 2013
norway$k600cc <- with(norway, 2.07+0.25*wind_mean^1.7) # lake area in km2. Cole & Caraco 1998
norway$k600mci <- with(norway, 2+1.5*wind_mean) # @MacIntyre2020

norway$ECO2vp <- with(norway, k600vp*(600/Sc)^1/2*24*1e-2*(pCO2 - pCO2.atm.2019)*K0*1e3*12.011) # gC/m2/day
norway$ECO2cc <- with(norway, k600cc*(600/Sc)^1/2*24*1e-2*(pCO2 - pCO2.atm.2019)*K0*1e3*12.011) # gC/m2/day
norway$ECO2mci <- with(norway, k600mci*(600/Sc)^1/2*24*1e-2*(pCO2 - pCO2.atm.2019)*K0*1e3*12.011) # gC/m2/day

norway$k600vp_mcmean <- NA
norway$k600vp_mcsd <- NA
norway$k600cc_mcmean <- NA
norway$k600cc_mcsd <- NA
norway$k600mci_mcmean <- NA
norway$k600mci_mcsd <- NA


for(i in 1:dim(norway)[1]){
  lake_area_i <- norway$lake_area_km2[i]
  Sc_i <- norway$Sc[i]
  pCO2_atm_i <- norway$pCO2_atm[i]
  pCO2_i <- norway$pCO2[i]
  K0_i <- norway$K0[i]
  
  pCO2_pred_i <- norway$pCO2_pred[i]
  pCO2_pred.sd_i <- norway$pCO2_pred.se[i]
  
  windm_i <- norway$wind_mean[i]
  windsd_i <- norway$wind_sd[i]
  wind <- runif(100000, windm_i - windsd_i, windm_i + windsd_i) 
  
  
  # k600 from Vachon & Prairie
  
  k600vp <- 2.51+1.48*wind+0.39*wind*log10(lake_area_i)
  k600vp_mi <- mean(k600vp)
  k600vp_sdi <- sd(k600vp)
  
  norway$k600vp_mcmean[i] <- k600vp_mi
  norway$k600vp_mcsd[i] <- k600vp_sdi

  
  ECO2vp <- k600vp*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2_atm_i)*K0_i*1e3*12.011 # gC/m2/day
  ECO2vp_mi <- mean(ECO2vp)
  ECO2vp_sdi <- sd(ECO2vp)
  
  norway$ECO2vp_mcmean[i] <- ECO2vp_mi
  norway$ECO2vp_mcsd[i] <- ECO2vp_sdi
  
  ECO2vp_pred <- k600vp*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_pred_i - pCO2_atm_i)*K0_i*1e3*12.011 # gC/m2/day
  ECO2vp_pred_mi <- mean(ECO2vp_pred)
  ECO2vp_pred_sdi <- sd(ECO2vp_pred)
  
  norway$ECO2vp_pred_mcmean[i] <- ECO2vp_pred_mi
  norway$ECO2vp_pred_mcsd[i] <- ECO2vp_pred_sdi
  
  
  # k600 from Cole and Caraco
  
  k600cc <- 2.07+0.25*wind^1.7
  k600cc_mi <- mean(k600cc)
  k600cc_sdi <- sd(k600cc)

  norway$k600cc_mcmean[i] <- k600cc_mi
  norway$k600cc_mcsd[i] <- k600cc_sdi
  
  ECO2cc <- k600cc*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2cc_mi <- mean(ECO2cc)
  ECO2cc_sdi <- sd(ECO2cc)
  
  norway$ECO2cc_mcmean[i] <- ECO2cc_mi
  norway$ECO2cc_mcsd[i] <- ECO2cc_sdi
  
  ECO2cc_pred <- k600cc*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_pred_i - pCO2_atm_i)*K0_i*1e3*12.011 # gC/m2/day
  ECO2cc_pred_mi <- mean(ECO2cc_pred)
  ECO2cc_pred_sdi <- sd(ECO2cc_pred)
  
  norway$ECO2cc_pred_mcmean[i] <- ECO2cc_pred_mi
  norway$ECO2cc_pred_mcsd[i] <- ECO2cc_pred_sdi
  
  # k600 from MacIntyre
  
  k600mci <- 2+1.5*wind
  k600mci_mi <- mean(k600mci)
  k600mci_sdi <- sd(k600mci)
  
  norway$k600mci_mcmean[i] <- k600mci_mi
  norway$k600mci_mcsd[i] <- k600mci_sdi
  
  ECO2mci <- k600mci*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_i - pCO2.atm.2019)*K0_i*1e3*12.011 # gC/m2/day
  ECO2mci_mi <- mean(ECO2mci)
  ECO2mci_sdi <- sd(ECO2mci)
  
  norway$ECO2mci_mcmean[i] <- ECO2mci_mi
  norway$ECO2mci_mcsd[i] <- ECO2mci_sdi
  
  ECO2mci_pred <- k600mci*(600/Sc_i)^(1/2)*24*1e-2*(pCO2_pred_i - pCO2_atm_i)*K0_i*1e3*12.011 # gC/m2/day
  ECO2mci_pred_mi <- mean(ECO2mci_pred)
  ECO2mci_pred_sdi <- sd(ECO2mci_pred)
  
  norway$ECO2mci_pred_mcmean[i] <- ECO2mci_pred_mi
  norway$ECO2mci_pred_mcsd[i] <- ECO2mci_pred_sdi
  
}


ggplot(norway)+
  geom_point(aes(y = ECO2vp_mcmean, x = as.character(Lake_ID), col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, x = as.character(Lake_ID), col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, x = as.character(Lake_ID), col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, x = as.character(Lake_ID), col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, x = as.character(Lake_ID), col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, x = as.character(Lake_ID), col = "MacIntyre"))+
  theme_minimal()
  
ggplot(norway, aes(x = TOC))+
  geom_point(aes(y = ECO2vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, col = "MacIntyre"))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "TOC, mg/L", y = "CO2 evasion in gC/m2/day", col = "")+
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()

ggplot(norway, aes(x = lake_area_km2))+
  geom_point(aes(y = ECO2vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, col = "MacIntyre"))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "Lake area, km2", y = "CO2 evasion in gC/m2/day", col = "")+
  scale_x_log10()+
  theme_minimal()


ggplot(norway, aes(x = ECO2cc_mcmean, y = ECO2mci_mcmean))+
  geom_point()+geom_abline(slope = 1, intercept = 0)

ggplot(norway)+geom_boxplot(aes(x = "ECO2cc_mcmean", y = ECO2cc_mcmean))+
  geom_boxplot(aes(x = "ECO2cc_pred_mcmean", y = ECO2cc_pred_mcmean))+
  geom_boxplot(aes(x = "ECO2vp_mcmean", y = ECO2vp_mcmean))+
  geom_boxplot(aes(x = "ECO2vp_pred_mcmean", y = ECO2vp_pred_mcmean))+
  geom_boxplot(aes(x = "ECO2mci_mcmean", y = ECO2mci_mcmean))+
  geom_boxplot(aes(x = "ECO2mci_pred_mcmean", y = ECO2mci_pred_mcmean))


ggplot(norway)+geom_point(aes(x = ECO2cc_mcmean, y = ECO2cc_pred_mcmean))+
  geom_text(x = 0.25, y = 0.05,label = paste0('r2 = ', round((cor(norway$ECO2cc_mcmean,norway$ECO2cc_pred_mcmean))^2,2)))+
  theme_minimal()

ggplot(norway)+geom_point(aes(x = ECO2vp_mcmean, y = ECO2vp_pred_mcmean))+
  geom_text(x = 1, y= 1,label = paste0('r2 = ', round((cor(norway$ECO2vp_mcmean,norway$ECO2vp_pred_mcmean))^2,2)))+
  theme_minimal()

ggplot(norway)+geom_point(aes(x = ECO2mci_mcmean, y = ECO2mci_pred_mcmean))+
  geom_text(x = 0.25, y= 0.05,label = paste0('r2 = ', round((cor(norway$ECO2mci_mcmean,norway$ECO2mci_pred_mcmean))^2,2)))+
  theme_minimal()

```

```{r figure-pCO2-toc}
# figure 2

ggplot(norway,aes(x = TOC))+geom_point(aes(y = pCO2_pred*1e6, col = "pCO2 predicted with TOC"), alpha = 0.5)+
  geom_point(aes(y = ta_pCO2*1e6, col = "pCO2 from TA"), alpha = 0.5)+
  geom_point(aes(y = pCO2*1e6, col = "Measured pCO2"), alpha = 0.5)+
  labs(y = "pCO2, μatm", x = "TOC, mg/L", col = "")+
  scale_color_manual(values = c( "firebrick","dodgerblue3","orange"))+
  theme_bw(base_size = 15)+
  theme(legend.position = "bottom")

ggplot(norway,aes(x = TOC_mol.L))+geom_point(aes(y = pCO2_pred, col = "pCO2 predicted with TOC"), alpha = 0.5)+
  geom_point(aes(y = ta_pCO2, col = "pCO2 from TA"), alpha = 0.5)+
  geom_point(aes(y = pCO2, col = "Measured pCO2"), alpha = 0.5)+
  labs(y = "pCO2, atm", x = "TOC, mol/L", col = "")+
  scale_color_manual(values = c( "firebrick","dodgerblue3","orange"))+
  annotate(geom = "label", x = 0.0014, y =0.03, 
           label = paste0("pCO2 (atm) = pCO2 atmosphere (atm) + ", pco2.toc.nor.sum$Estimate, "*TOC (mol/L) ± ", pco2.toc.nor.sum$`Std. Error`), col = "orange")+  
  theme_bw(base_size = 15)+
  theme(legend.position = "bottom")

```

# NLS

```{r import-nls-data}

ion <- readRDS("ion.rds")
lakes <- readRDS("lakes.rds")

nls0 <- merge(ion, lakes, by = "ebint")

nls0$uncertain <- ifelse(nls0$dist_closest_ebint/nls0$dist_2nd_closest_ebint > 0.5, FALSE, TRUE)
nls0$uncertain[which(is.na(nls0$uncertain)==TRUE)] <- FALSE

nls0$temp_c[which(is.na(nls0$temp_c)==T)] <- 4
nls0$lake_area_km2 <- nls0$area * 1e-6 #m2 to km2

duplicated.id <- nls0$ebint[which(duplicated(nls0$ebint) ==T)]

nls <- nls0 %>% filter(toc >= 0) %>% filter(tp > 0) %>% filter(uncertain == FALSE) %>%
  dplyr::select(c("ebint", "longitude", "latitude","nation", "temp_c", "toc", "ph", "alk_ueq","lake_area_km2")) %>%
  rename(Lake_ID = ebint, Long = longitude, Lat = latitude, TOC = toc, pH = ph, TA = alk_ueq)

nls$TOC_mol.L <- with(nls, TOC/12.011*1e-3)
nls$TA_mol.L <- nls$TA*1e-6 # convert from ueq/L to mol/L
nls$pCO2_atm <- pCO2.atm.1995
nls$temp_K <- with(nls, temp_c + 273.15)

nls.sf <- nls %>% group_by(Lake_ID, nation) %>% summarise_all(median) %>% st_as_sf(coords = c("Long","Lat"))

nls.sf$Kw <- with(nls.sf, exp(148.9802 - 13847.26/temp_K - 23.6521*log(temp_K))) # Dickson and Riley, 1979

nls.sf$K0 <- with(nls.sf, exp(-58.0931+90.5069*100/temp_K+22.2940*log(temp_K/100))) # Weiss 1974
nls.sf$K1 <- with(nls.sf, 10^-((3404.71/temp_K)- 14.8435 + 0.032786*temp_K)) # Harned and Davis, 1943
nls.sf$K2 <- with(nls.sf, 10^-((2902.39/temp_K) - 6.4980 + 0.02379*temp_K)) # Harned and Scholes, 1941

nls.sf$Hplus <- 10^(-nls.sf$pH)
nls.sf$OH <- with(nls.sf, Kw/Hplus)

nls.sf$ta_CO3 <- with(nls.sf,(TA_mol.L - OH + Hplus)/(Hplus/K2 +2)) # in molL
nls.sf$ta_HCO3 <- with(nls.sf, ta_CO3*Hplus/K2)
nls.sf$ta_H2CO3 <- with(nls.sf, ta_HCO3*Hplus/K1)
nls.sf$ta_pCO2 <- with(nls.sf,ta_H2CO3/K0)

nls.sf$ta_DIC <- with(nls.sf, ta_H2CO3 + ta_HCO3 + ta_CO3)

```

```{r wind-speed-10-years-mean-nls, eval = T}
sd.function <- function(x) {sqrt(sum(x)/120)}

u10.1995 <- raster::stack("ERA5/u10_1990_1999.nc")
u10.1995.mean <- u10.1995 %>% mean(na.rm = T)
u10.1995.sd <- (u10.1995 - u10.1995.mean)^2 %>% stack %>% sd.function()

v10.1995 <- raster::stack("ERA5/v10_1990_1999.nc")
v10.1995.mean <- v10.1995 %>% mean(na.rm = T)
v10.1995.sd <- (v10.1995 - v10.1995.mean)^2 %>% stack %>% sd.function()

wind.1995.mean <- sqrt(u10.1995^2 + v10.1995^2) %>% mean()
wind.1995.sd <- (sqrt(u10.1995^2 + v10.1995^2) - wind.1995.mean)^2 %>% stack %>% sd.function()

wind.1995.nls <- raster::extract(wind.1995.mean, nls.sf)
nls.sf$wind_mean <- wind.1995.nls
#wind.1995.nls.df <- wind.1995.nls@data %>% setNames(c("Long","Lat","Lake_ID","wind_mean"))

wind.1995.sd.nls <- raster::extract(wind.1995.sd, nls.sf)
nls.sf$wind_sd <- wind.1995.sd.nls
#wind.1995.sd.nls.df <- wind.1995.sd.nls@data %>% setNames(c("Long","Lat","Lake_ID","wind_sd"))

# nls.sf <- nls %>% merge(wind.1995.nls.df, by = intersect(names(nls), names(wind.1995.nls.df)))%>%
#   merge(wind.1995.sd.nls.df, by = intersect(names(nls), names(wind.1995.nls.df)))
```

```{r k600-monte-carlo-nls}
nls.sf$Sc <- with(nls.sf, 1923.6-125.06*temp_c+4.3733*temp_c^2-0.086781*temp_c^3)


nls.sf$pCO2 <- predict(pco2.toc.nor, newdata = nls.sf, type = "response")
nls.sf$pCO2.se <- predict(pco2.toc.nor, newdata = nls.sf, type = "response", se.fit = T)$se.fit

nls.sf$k600vp <- with(nls.sf, 2.51+1.48*wind_mean+0.39*wind_mean*log10(lake_area_km2)) # lake.area already in km2, in cm/h. Vachon and Prairie 2013
nls.sf$k600cc <- with(nls.sf, 2.07+0.25*wind_mean^1.7) # lake area in km2. Cole & Caraco 1998
nls.sf$k600mci <- with(nls.sf, 2+1.5*wind_mean) # @MacIntyre2020

nls.sf$ECO2vp <- with(nls.sf, k600vp*(600/Sc)^(1/2)*24*1e-2*(pCO2 - pCO2.atm.2019)*K0*1e3*12.011) # gC/m2/day
nls.sf$ECO2cc <- with(nls.sf, k600cc*(600/Sc)^(1/2)*24*1e-2*(pCO2 - pCO2.atm.2019)*K0*1e3*12.011) # gC/m2/day
nls.sf$ECO2mci <- with(nls.sf, k600mci*(600/Sc)^(1/2)*24*1e-2*(pCO2 - pCO2.atm.2019)*K0*1e3*12.011) # gC/m2/day

nls.sf$k600vp_mcmean <- NA
nls.sf$k600vp_mcsd <- NA
nls.sf$k600cc_mcmean <- NA
nls.sf$k600cc_mcsd <- NA
nls.sf$k600mci_mcmean <- NA
nls.sf$k600mci_mcsd <- NA


for(i in 1:dim(nls.sf)[1]){
  lake_area_i <- nls.sf$lake_area_km2[i]
  Sc_i <- nls.sf$Sc[i]
  K0_i <- nls.sf$K0[i]

  
  windm_i <- nls.sf$wind_mean[i]
  windsd_i <- nls.sf$wind_sd[i]
  wind <- runif(100000, windm_i - windsd_i, windm_i + windsd_i) 
  
  pCO2_i <- nls.sf$pCO2[i]
  pCO2se_i <- nls.sf$pCO2.se[i]
  pCO2 <-  runif(100000, pCO2_i - pCO2se_i, pCO2_i + pCO2se_i) 
  
  # k600 from Vachon & Prairie
  
  k600vp <- 2.51+1.48*wind+0.39*wind*log10(lake_area_i)
  k600vp_mi <- mean(k600vp)
  k600vp_sdi <- sd(k600vp)
  
  nls.sf$k600vp_mcmean[i] <- k600vp_mi
  nls.sf$k600vp_mcsd[i] <- k600vp_sdi

  
  ECO2vp <- k600vp*(600/Sc_i)^(1/2)*24*1e-2*(pCO2 - pCO2.atm.1995)*K0_i*1e3*12.011 # gC/m2/day
  ECO2vp_mi <- mean(ECO2vp)
  ECO2vp_sdi <- sd(ECO2vp)
  
  nls.sf$ECO2vp_mcmean[i] <- ECO2vp_mi
  nls.sf$ECO2vp_mcsd[i] <- ECO2vp_sdi
  
  # k600 from Cole and Caraco
  
  k600cc <- 2.07+0.25*wind^1.7
  k600cc_mi <- mean(k600cc)
  k600cc_sdi <- sd(k600cc)

  nls.sf$k600cc_mcmean[i] <- k600cc_mi
  nls.sf$k600cc_mcsd[i] <- k600cc_sdi
  
  ECO2cc <- k600cc*(600/Sc_i)^(1/2)*24*1e-2*(pCO2 - pCO2.atm.1995)*K0_i*1e3*12.011 # gC/m2/day
  ECO2cc_mi <- mean(ECO2cc)
  ECO2cc_sdi <- sd(ECO2cc)
  
  nls.sf$ECO2cc_mcmean[i] <- ECO2cc_mi
  nls.sf$ECO2cc_mcsd[i] <- ECO2cc_sdi
  
  # k600 from MacIntyre
  
  k600mci <- 2+1.5*wind
  k600mci_mi <- mean(k600mci)
  k600mci_sdi <- sd(k600mci)
  
  nls.sf$k600mci_mcmean[i] <- k600mci_mi
  nls.sf$k600mci_mcsd[i] <- k600mci_sdi
  
  ECO2mci <- k600mci*(600/Sc_i)^(1/2)*24*1e-2*(pCO2 - pCO2.atm.1995)*K0_i*1e3*12.011 # gC/m2/day
  ECO2mci_mi <- mean(ECO2mci)
  ECO2mci_sdi <- sd(ECO2mci)
  
  nls.sf$ECO2mci_mcmean[i] <- ECO2mci_mi
  nls.sf$ECO2mci_mcsd[i] <- ECO2mci_sdi
  
}

  
ggplot(nls.sf, aes(x = TOC))+
  geom_point(aes(y = ECO2vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = ECO2vp_mcmean - ECO2vp_mcsd, ymax = ECO2vp_mcmean + ECO2vp_mcsd, col = "Vachon & Prairie"))+
  geom_point(aes(y = ECO2cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = ECO2cc_mcmean - ECO2cc_mcsd, ymax = ECO2cc_mcmean + ECO2cc_mcsd, col = "Cole & Caraco"))+
  geom_point(aes(y = ECO2mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = ECO2mci_mcmean - ECO2mci_mcsd, ymax = ECO2mci_mcmean + ECO2mci_mcsd, col = "MacIntyre"))+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "TOC, mg/L", y = "CO2 evasion in gC/m2/day", col = "")+
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()
```

```{r view-k600}

ggplot(nls.sf, aes(x = lake_area_km2))+
  geom_point(aes(y = k600vp_mcmean, col = "Vachon & Prairie"))+
  geom_errorbar(aes(ymin = k600vp_mcmean - k600vp_mcsd, ymax = k600vp_mcmean + k600vp_mcsd, col = "Vachon & Prairie"), alpha = 0.5)+
  geom_point(aes(y = k600cc_mcmean, col = "Cole & Caraco"))+
  geom_errorbar(aes(ymin = k600cc_mcmean - k600cc_mcsd, ymax = k600cc_mcmean + k600cc_mcsd, col = "Cole & Caraco"), alpha = 0.5)+
  geom_point(aes(y = k600mci_mcmean, col = "MacIntyre"))+
  geom_errorbar(aes(ymin = k600mci_mcmean - k600mci_mcsd, ymax = k600mci_mcmean + k600mci_mcsd, col = "MacIntyre"), alpha = 0.5)+
  scale_color_manual(values = c("orange","firebrick","dodgerblue3"))+
  labs(x = "Lake area, km2", y = "k600 in cm/h", col = "")+
  scale_x_log10()+
  theme_minimal()


```
```{r compare-k600-nls-norway}
ggarrange(
ggplot(nls.sf)+
  geom_boxplot(aes(x = "Cole and Caraco", y = k600cc_mcmean), fill = "orange")+
  geom_boxplot(aes(x = "MacIntyre", y = k600mci_mcmean), fill = "firebrick")+
  geom_boxplot(aes(x = "Vachon and Prairie", y = k600vp_mcmean), fill = "dodgerblue3")+
  labs(x="", y = "k600, cm/h", title = "Northern Lake Survey")+
  lims(y = c(2,8))+
  theme_minimal(),
ggplot(norway)+
  geom_boxplot(aes(x = "Cole and Caraco", y = k600cc_mcmean), fill = "orange")+
  geom_boxplot(aes(x = "MacIntyre", y = k600mci_mcmean), fill = "firebrick")+
  geom_boxplot(aes(x = "Vachon and Prairie", y = k600vp_mcmean), fill = "dodgerblue3")+
  labs(x="", y = "k600, cm/h", title = "CBA + N112 survey")+
    lims(y = c(2,8))+
  theme_minimal(),
ncol = 2
)

kable(data.frame(
  range = c(paste(round(range(norway$k600cc_mcmean),1), collapse = " to "), paste(round(range(norway$k600mci_mcmean),1), collapse = " to "), paste(round(range(norway$k600vp_mcmean),1), collapse = " to "), paste(round(range(nls.sf$k600cc_mcmean),1), collapse = " to "), paste(round(range(nls.sf$k600mci_mcmean),1), collapse = " to "), paste(round(range(nls.sf$k600vp_mcmean),1), collapse = " to ")),
                 mean = c(round(mean(norway$k600cc_mcmean),1), round(mean(norway$k600mci_mcmean),1), round(mean(norway$k600vp_mcmean),1), round(mean(nls.sf$k600cc_mcmean),1), round(mean(nls.sf$k600mci_mcmean),1), round(mean(nls.sf$k600vp_mcmean),1)))) %>%
  kableExtra::group_rows(group_label = "CBA and N112", start_row = 1, end_row = 3) %>%
  kableExtra::group_rows(group_label = "Northern Lakes Survey", start_row = 4, end_row = 6)


```

# Compare ECO2 in the 3 surveys

```{r summary-eco2-montecarlo, results = T}
no <- norway %>% 
  dplyr::select(c("survey", "pCO2" , "ECO2cc_mcmean", "ECO2mci_mcmean", "ECO2vp_mcmean","ECO2cc_pred_mcmean",
                  "ECO2mci_pred_mcmean","ECO2vp_pred_mcmean")) %>%
  st_drop_geometry()

medyear <- function(x){median(x, na.rm = T) * 365} # to convert gC/m2/d in gC/m2/year
avyear <- function(x){mean(x, na.rm = T) * 365}
sdyear<- function(x){sd(x, na.rm = T) * 365}

summary_no.1 <- no %>% group_by(survey) %>% summarize_all(medyear) %>% mutate(pCO2 = pCO2/365*1e6) #divide fCO2 by 365 to come back to mean concentration value
summary_no.2 <- no %>% group_by(survey) %>% summarize_all(avyear) %>% mutate(pCO2 = pCO2/365*1e6) # + convert to uatm
summary_no.3 <- no %>% group_by(survey) %>% summarize_all(sdyear) %>% mutate(pCO2 = pCO2/365*1e6)

no2 <- rbind(summary_no.1, summary_no.2, summary_no.3) %>% arrange(survey) %>% 
  mutate(stat = c("median","mean","standard deviation","median","mean","standard deviation")) %>%
  dplyr::select(c("survey", "stat" ,"pCO2", "ECO2cc_mcmean", "ECO2cc_pred_mcmean", "ECO2mci_mcmean", "ECO2mci_pred_mcmean", "ECO2vp_mcmean", "ECO2vp_pred_mcmean")) 

fen <- nls.sf  %>%
  st_drop_geometry() %>% 
  dplyr::select(c("pCO2" , "ECO2cc_mcmean", "ECO2mci_mcmean","ECO2vp_mcmean")) %>% 
  setNames(c("Lake_ID","pCO2","ECO2cc_pred_mcmean",  "ECO2mci_pred_mcmean","ECO2vp_pred_mcmean")) %>% 
  mutate(ECO2cc_mcmean = NA, ECO2mci_mcmean = NA, ECO2vp_mcmean = NA, pCO2 = pCO2/365*1e6, survey = "NLS") %>%
  dplyr::select(c("survey","pCO2", "ECO2cc_mcmean", "ECO2mci_mcmean", "ECO2vp_mcmean", "ECO2cc_pred_mcmean", "ECO2mci_pred_mcmean", "ECO2vp_pred_mcmean"))



summary_fen.1 <- fen %>% group_by(survey) %>% summarize_all(medyear) %>% mutate(stat = "median")
summary_fen.2 <- fen %>% group_by(survey) %>% summarize_all(avyear) %>% mutate(stat = "mean")
summary_fen.3 <- fen %>% group_by(survey) %>% summarize_all(sdyear) %>% mutate(stat = "standard deviation")

fen2 <- rbind(summary_fen.1, summary_fen.2, summary_fen.3) %>% mutate(stat = c("median","mean","standard deviation")) %>%
    dplyr::select(c("survey", "stat" ,"pCO2",
                    "ECO2cc_mcmean", "ECO2cc_pred_mcmean",
                    "ECO2mci_mcmean", "ECO2mci_pred_mcmean",
                    "ECO2vp_mcmean", "ECO2vp_pred_mcmean")) 

tot <- rbind(no2, fen2)

kable(tot[,-1], digits = 2, col.names = c(" ","pCO2, uatm", "ECO2 in gC/m2/year","ECO2 from predicted pCO2 in gC/m2/year","ECO2 in gC/m2/year","ECO2 from predicted pCO2 in gC/m2/year", "ECO2 in gC/m2/year","ECO2 from predicted pCO2 in gC/m2/year")) %>% 
  kable_styling %>% 
  kableExtra::group_rows(group_label= "CBA", start_row = 1, end_row = 3) %>%
  kableExtra::group_rows(group_label= "N112", start_row = 4, end_row = 6) %>% 
  kableExtra::group_rows(group_label= "NLS", start_row = 7, end_row = 9) %>% 
  kableExtra::add_header_above(c(" " = 2, "Cole and Caraco" = 2, "MacIntyre" = 2, "Vachon and Prairie" = 2)) 

# kable(tot, digits = 2, col.names = c(" ","pCO2, uatm", "ECO2 from DIC in gC/m2/year","ECO2 from TOC in gC/m2/year","ECO2 from DIC in gC/m2/year","ECO2 from TOC in gC/m2/year")) %>% 
#   kable_styling() %>% 
#   kableExtra::group_rows(group_label= "CBA", start_row = 1, end_row = 3) %>%
#   kableExtra::group_rows(group_label= "N112", start_row = 4, end_row = 6) %>%
#   kableExtra::group_rows(group_label= "NLS", start_row = 7, end_row = 9) %>% 
#   kableExtra::add_header_above(c(" " = 2, "High estimates" = 2, "Low estimates" = 2)) %>%
#   save_kable(file = "compare.eco2.measured.png",zoom = 5)

```


```{r plot-pCO2-pred-nls}

g1 <- ggplot(data = nls.sf, aes(x = pH))+
  geom_point(aes(y = ta_pCO2*1e6, col = "pCO2 from carbonate equilibrium"), alpha = 0.5)+
  geom_point(aes(y = pCO2*1e6, col = "pCO2 modelled with TOC"), alpha = 0.5)+
  geom_hline(yintercept = pCO2.atm.1995, lty = 2)+
  labs(x = "pH", col = "", y = "pCO2, μatm")+
  scale_color_manual(values = c("firebrick","dodgerblue"))+
  theme_bw(base_size = 15)+
  theme(legend.position = "top")+
  facet_zoom(y = ta_pCO2 > 0, zoom.size = 2)


g2 <- ggplot(data = nls.sf, aes(x = pH))+
  geom_point(aes(y = ta_pCO2*1e6, col = "pCO2 from carbonate equilibrium"), alpha = 0.5)+
  geom_point(aes(y = pCO2*1e6, col = "pCO2 modelled with TOC"), alpha = 0.5)+
  geom_hline(yintercept = pCO2.atm.1995, lty = 2)+
  labs(x = "pH", col = "", y = "pCO2, μatm")+
  scale_color_manual(values = c("firebrick","dodgerblue"))+
  ylim(0,0.006*1e6)+
  theme_bw(base_size = 15)+
  theme(legend.position = "top")

ggarrange(g2,g1, common.legend = T)
```


# FCO2 in Fennoscandia

```{r emission-per-unit-area-mc}
inland.waters.nor <- 20221 * 1e6 # km2 to m2, @SSB
land.nor <- land <- 323806 * 1e6 # km2 to m2, @SSB
inland.waters.swe <- 4014169.92 * 1e4 # ha to m2, @SCB
land.swe <- 444435 * 1e6 # km2 to m2
inland.waters.fin <- 34515 * 1e6 # km2 to m2, @Statfi
land.fin <- 338478 * 1e6 # km2 to m2

tot.water <- inland.waters.nor + inland.waters.swe + inland.waters.fin
tot.land <- land.nor + land.swe + land.fin


```


```{r emission-per-unit-land}
inland.waters.nor <- 20221 * 1e6 # km2 to m2, @SSB
land.nor <- land <- 323806 * 1e6 # km2 to m2, @SSB
inland.waters.swe <- 4014169.92 * 1e4 # ha to m2, @SCB
land.swe <- 444435 * 1e6 # km2 to m2
inland.waters.fin <- 34515 * 1e6 # km2 to m2, @Statfi
land.fin <- 338478 * 1e6 # km2 to m2

tot.water <- inland.waters.nor + inland.waters.swe + inland.waters.fin
tot.land <- land.nor + land.swe + land.fin

library(DescTools)

median.eco2.cc <- c(unname(MedianCI(nls.sf$ECO2cc_mcmean, na.rm = T)[1])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[1])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[1])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[1])*365))

median.eco2.cc.ciinf <- c(unname(MedianCI(nls.sf$ECO2cc_mcmean, na.rm = T)[2])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[2])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[2])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[2])*365))

median.eco2.cc.cisup <- c(unname(MedianCI(nls.sf$ECO2cc_mcmean, na.rm = T)[3])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[3])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[3])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2cc_mcmean, na.rm = T)[3])*365))



median.eco2.mci <- c(unname(MedianCI(nls.sf$ECO2mci_mcmean, na.rm = T)[1])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[1])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[1])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[1])*365))

median.eco2.mci.ciinf <- c(unname(MedianCI(nls.sf$ECO2mci_mcmean, na.rm = T)[2])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[2])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[2])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[2])*365))

median.eco2.mci.cisup <- c(unname(MedianCI(nls.sf$ECO2mci_mcmean, na.rm = T)[3])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[3])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[3])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2mci_mcmean, na.rm = T)[3])*365))



median.eco2.vp <- c(unname(MedianCI(nls.sf$ECO2vp_mcmean, na.rm = T)[1])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[1])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[1])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[1])*365))

median.eco2.vp.ciinf <- c(unname(MedianCI(nls.sf$ECO2mci_mcmean, na.rm = T)[2])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[2])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[2])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[2])*365))

median.eco2.vp.cisup <- c(unname(MedianCI(nls.sf$ECO2vp_mcmean, na.rm = T)[3])*365, #NLSS # in gC/m2/year
                    with(subset(nls.sf, nation == "Norway"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[3])*365),
                    with(subset(nls.sf, nation == "Sweden"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[3])*365),
                    with(subset(nls.sf, nation == "Finland"), unname(MedianCI(ECO2vp_mcmean, na.rm = T)[3])*365))


land.eco2.cc <- c(median.eco2.cc[1]*tot.water/tot.land,
                    median.eco2.cc[2]*inland.waters.nor/land.nor,
                    median.eco2.cc[3]*inland.waters.swe/land.swe,
                    median.eco2.cc[4]*inland.waters.fin/land.fin)

land.eco2.cc.ciinf <- c(median.eco2.cc.ciinf[1]*tot.water/tot.land,
                    median.eco2.cc.ciinf[2]*inland.waters.nor/land.nor,
                    median.eco2.cc.ciinf[3]*inland.waters.swe/land.swe,
                    median.eco2.cc.ciinf[4]*inland.waters.fin/land.fin)

land.eco2.cc.cisup <- c(median.eco2.cc.cisup[1]*tot.water/tot.land,
                    median.eco2.cc.cisup[2]*inland.waters.nor/land.nor,
                    median.eco2.cc.cisup[3]*inland.waters.swe/land.swe,
                    median.eco2.cc.cisup[4]*inland.waters.fin/land.fin)

land.eco2.mci <- c(median.eco2.mci[1]*tot.water/tot.land,
                    median.eco2.mci[2]*inland.waters.nor/land.nor,
                    median.eco2.mci[3]*inland.waters.swe/land.swe,
                    median.eco2.mci[4]*inland.waters.fin/land.fin)

land.eco2.mci.ciinf <- c(median.eco2.mci.ciinf[1]*tot.water/tot.land,
                    median.eco2.mci.ciinf[2]*inland.waters.nor/land.nor,
                    median.eco2.mci.ciinf[3]*inland.waters.swe/land.swe,
                    median.eco2.mci.ciinf[4]*inland.waters.fin/land.fin)

land.eco2.mci.cisup <- c(median.eco2.mci.cisup[1]*tot.water/tot.land,
                    median.eco2.mci.cisup[2]*inland.waters.nor/land.nor,
                    median.eco2.mci.cisup[3]*inland.waters.swe/land.swe,
                    median.eco2.mci.cisup[4]*inland.waters.fin/land.fin)

land.eco2.vp <- c(median.eco2.vp[1]*tot.water/tot.land,
                    median.eco2.vp[2]*inland.waters.nor/land.nor,
                    median.eco2.vp[3]*inland.waters.swe/land.swe,
                    median.eco2.vp[4]*inland.waters.fin/land.fin)

land.eco2.vp.ciinf <- c(median.eco2.vp.ciinf[1]*tot.water/tot.land,
                    median.eco2.vp.ciinf[2]*inland.waters.nor/land.nor,
                    median.eco2.vp.ciinf[3]*inland.waters.swe/land.swe,
                    median.eco2.vp.ciinf[4]*inland.waters.fin/land.fin)

land.eco2.vp.cisup <- c(median.eco2.vp.cisup[1]*tot.water/tot.land,
                    median.eco2.vp.cisup[2]*inland.waters.nor/land.nor,
                    median.eco2.vp.cisup[3]*inland.waters.swe/land.swe,
                    median.eco2.vp.cisup[4]*inland.waters.fin/land.fin)

kable(tibble("Country" = c("All 3", "Norway", "Sweden", "Finland"),
            "Water"= c(tot.water, inland.waters.nor, inland.waters.swe, inland.waters.fin),
            "Land" = c(tot.land, land.nor, land.swe, land.fin),
            median.eco2.cc,
            median.eco2.cc.ciinf,
            median.eco2.cc.cisup,
            land.eco2.cc,
            land.eco2.cc.ciinf,
            land.eco2.cc.cisup,
            median.eco2.mci,
            median.eco2.mci.ciinf,
            median.eco2.mci.cisup,
            land.eco2.mci,
            land.eco2.mci.ciinf,
            land.eco2.mci.cisup,
            median.eco2.vp,
            median.eco2.vp.ciinf,
            median.eco2.vp.cisup,
            land.eco2.vp,
            land.eco2.vp.ciinf,
            land.eco2.vp.cisup),
      col.names = c("Country","Water area, km2", "Land area, km2", "Median ECO2, gC/m2/year","CI inf", "CI sup", "Median ECO2, gC/m2 land/year","CI inf", "CI sup", "Median ECO2, gC/m2/year","CI inf", "CI sup", "Median ECO2, gC/m2 land/year","CI inf", "CI sup","Median ECO2, gC/m2/year","CI inf", "CI sup", "Median ECO2, gC/m2 land/year","CI inf", "CI sup"),
      digits = 2) %>% kable_styling() %>%
  add_header_above(c(" ","Area" = 2, "Cole and Caraco" = 6, "MacIntyre" = 6, "Vachon and Prairie" = 6))

```

```{r bin, results = T}
breaks.lakes <- c(0,0.1,1,10,100,Inf)

nls.sf$lake.category <- cut(nls.sf$lake_area_km2, breaks.lakes)

nls.sf$inland.water <- NA
nls.sf$inland.water[which(nls.sf$nation == "Norway")] <- 20221  # km2 , @SSB
nls.sf$inland.water[which(nls.sf$nation == "Sweden")] <- 4014169.92 * 1e-2 # ha to km2, @SCB
nls.sf$inland.water[which(nls.sf$nation == "Finland")] <- 34515 # km2, @Statfi

nls.sf$lake.area.under100 <- NA
nls.sf$lake.area.under100[which(nls.sf$nation == "Norway")] <- sum(subset(nls.sf, nation == "Norway" & lake_area_km2 < 100)$lake_area_km2) # nls.sf area without biggest lakes
nls.sf$lake.area.under100[which(nls.sf$nation == "Sweden")] <- sum(subset(nls.sf, nation == "Sweden" & lake_area_km2 < 100)$lake_area_km2)# nls.sf area without biggest lakes
nls.sf$lake.area.under100[which(nls.sf$nation == "Finland")] <- sum(subset(nls.sf, nation == "Finland" & lake_area_km2 < 100)$lake_area_km2) # nls.sf area without biggest lakes

nls.sf$lake.area.over100 <- NA
nls.sf$lake.area.over100[which(nls.sf$nation == "Norway")] <- sum(subset(nls.sf, nation == "Norway" & lake_area_km2 > 100)$lake_area_km2) # areas of biggest lakes
nls.sf$lake.area.over100[which(nls.sf$nation == "Sweden")] <- sum(subset(nls.sf, nation == "Sweden" & lake_area_km2 > 100)$lake_area_km2) # areas of biggest lakes
nls.sf$lake.area.over100[which(nls.sf$nation == "Finland")] <- sum(subset(nls.sf, nation == "Finland" & lake_area_km2 > 100)$lake_area_km2) # areas of biggest lakes

# uncertainty on bins??? BOOTSTRAP???

binned.nls.sf <- nls.sf %>% st_drop_geometry() %>% group_by(nation, inland.water, lake.area.under100, lake.area.over100, lake.category) %>% 
                        summarise(nb.lakes = n(),
                                  cat.lake.area = sum(lake_area_km2),
                                  prop.cat = ifelse(lake.category != "(100,Inf]", cat.lake.area/lake.area.under100 * 100, NA),
                                  pCO2_mean = mean(pCO2 * 1e6, na.rm = T),  # in uatm
                                  pCO2_sd = sd(pCO2 * 1e6, na.rm = T),
                                  k600vp_mean = mean(k600vp_mcmean, na.rm = T),
                                  k600vp_sd = sd(k600vp_mcmean, na.rm = T), # dispersion and not uncertainty
                                  ECO2vp = mean(ECO2vp_mcmean * 365, na.rm = T), # in gC/m2/year
                                  ECO2vp_sd = sd(ECO2vp_mcmean * 365, na.rm = T),
                                  k600cc_mean = mean(k600cc_mcmean, na.rm = T), # in cm/h
                                  k600cc_sd = sd(k600cc_mcmean, na.rm = T),
                                  ECO2cc = mean(ECO2cc_mcmean * 365, na.rm = T),
                                  ECO2cc_sd = sd(ECO2cc_mcmean * 365, na.rm = T),
                                  k600mci_mean = mean(k600mci_mcmean, na.rm = T), # in cm/h
                                  k600mci_sd = sd(k600mci_mcmean, na.rm = T),
                                  ECO2mci = mean(ECO2mci_mcmean * 365, na.rm = T),
                                  ECO2mci_sd = sd(ECO2mci_mcmean * 365, na.rm = T)
                                  ) %>%
                          mutate(FCO2vp = ifelse(lake.category != "(100,Inf]",
                                               ECO2vp*(cat.lake.area/lake.area.under100*(inland.water-lake.area.over100))*1e6/12.011 * 43.99 * 1e-12, #MtCO2/y. ECO2 multiplied by the proportion of lakes belonging to each category in Norway, lake area *1e6 to convert in m2,
                                               ECO2vp*lake.area.over100*1e6/12.011 * 43.99 * 1e-12),
                                 FCO2cc = ifelse(lake.category != "(100,Inf]", 
                                                   ECO2cc*(cat.lake.area/lake.area.under100*(inland.water-lake.area.over100))*1e6/12.011 * 43.99 * 1e-12,
                                                   ECO2cc*lake.area.over100*1e6/12.011 * 43.99 * 1e-12),
                                 FCO2mci = ifelse(lake.category != "(100,Inf]", 
                                                   ECO2mci*(cat.lake.area/lake.area.under100*(inland.water-lake.area.over100))*1e6/12.011 * 43.99 * 1e-12,
                                                   ECO2mci*lake.area.over100*1e6/12.011 * 43.99 * 1e-12)
                                 ) %>%  #MtCO2/y
  unique()

# doing a monte carlo for uncertainty on FCO2 (using ECO2_sd for proparating error)
z <- 1.96 # for 95% confidence interval

for(i in 1:dim(binned.nls.sf)[1]){
 cat.lake.area_i <- binned.nls.sf$cat.lake.area[i]
 lake.area.under100_i <- binned.nls.sf$lake.area.under100[i]
 lake.area.over100_i <- binned.nls.sf$lake.area.over100[i]
 inland.water_i <- binned.nls.sf$inland.water[i]
 
 lake.category_i <- binned.nls.sf$lake.category[i]
 
 ECO2cc_i <- binned.nls.sf$ECO2cc[i]
 ECO2ccsd_i <- binned.nls.sf$ECO2cc_sd[i]
  ECO2cc <- runif(100000, ECO2cc_i - ECO2ccsd_i, ECO2cc_i + ECO2ccsd_i)
   
 ECO2mci_i <- binned.nls.sf$ECO2mci[i]
 ECO2mcisd_i <- binned.nls.sf$ECO2mci_sd[i]
 ECO2mci <- runif(100000, ECO2mci_i - ECO2mcisd_i, ECO2mci_i + ECO2mcisd_i)
 
  ECO2vp_i <- binned.nls.sf$ECO2vp[i]
 ECO2vpsd_i <- binned.nls.sf$ECO2vp_sd[i]
 ECO2vp <- runif(100000, ECO2vp_i - ECO2vpsd_i, ECO2vp_i + ECO2vpsd_i)
 
 if(lake.category_i != "(100,Inf]"){
   FCO2cc <- ECO2cc*(cat.lake.area_i/lake.area.under100_i*(inland.water_i-lake.area.over100_i))*1e6/12.011 * 43.99 * 1e-12
   FCO2mci <- ECO2mci*(cat.lake.area_i/lake.area.under100_i*(inland.water_i-lake.area.over100_i))*1e6/12.011 * 43.99 * 1e-12
  FCO2vp <- ECO2vp*(cat.lake.area_i/lake.area.under100_i*(inland.water_i-lake.area.over100_i))*1e6/12.011 * 43.99 * 1e-12 #MtCO2/y. ECO2 multiplied by the proportion of lakes belonging to each category in Norway, lake area *1e6 to convert in m2,

 }else if(lake.category_i == "(100,Inf]"){
   FCO2cc <- ECO2cc*lake.area.over100_i*1e6/12.011 * 43.99 * 1e-12
   FCO2mci <- ECO2mci*lake.area.over100_i*1e6/12.011 * 43.99 * 1e-12
   FCO2vp <- ECO2vp*lake.area.over100_i*1e6/12.011 * 43.99 * 1e-12
 }
 
 
  binned.nls.sf$FCO2cc_mcmean[i] <- mean(FCO2cc, na.rm = T)
  binned.nls.sf$FCO2mci_mcmean[i] <- mean(FCO2mci, na.rm = T)
  binned.nls.sf$FCO2vp_mcmean[i] <- mean(FCO2vp, na.rm = T)
  
  binned.nls.sf$FCO2cc_mcsd[i] <- sd(FCO2cc, na.rm = T)
  binned.nls.sf$FCO2mci_mcsd[i] <- sd(FCO2mci, na.rm = T)
  binned.nls.sf$FCO2vp_mcsd[i] <- sd(FCO2vp, na.rm = T)
  
  binned.nls.sf$FCO2cc_confintplus[i] <- mean(FCO2cc, na.rm = T) + z*sd(FCO2cc, na.rm = T)/sqrt(100000)
  binned.nls.sf$FCO2mci_confintplus[i] <- mean(FCO2mci, na.rm = T) + z*sd(FCO2mci, na.rm = T)/sqrt(100000)
  binned.nls.sf$FCO2vp_confintplus[i] <- mean(FCO2vp, na.rm = T) + z*sd(FCO2vp, na.rm = T)/sqrt(100000)
      
  binned.nls.sf$FCO2cc_confintminus[i] <- mean(FCO2cc, na.rm = T) - z*sd(FCO2cc, na.rm = T)/sqrt(100000)
  binned.nls.sf$FCO2mci_confintminus[i] <- mean(FCO2mci, na.rm = T) - z*sd(FCO2mci, na.rm = T)/sqrt(100000)
  binned.nls.sf$FCO2vp_confintminus[i] <- mean(FCO2vp, na.rm = T) - z*sd(FCO2vp, na.rm = T)/sqrt(100000)
  
    binned.nls.sf$FCO2cc_confintminus[i] <- mean(FCO2cc, na.rm = T) - z*sd(FCO2cc, na.rm = T)/sqrt(100000)
  binned.nls.sf$FCO2mci_confintminus[i] <- mean(FCO2mci, na.rm = T) - z*sd(FCO2mci, na.rm = T)/sqrt(100000)
  binned.nls.sf$FCO2vp_confintminus[i] <- mean(FCO2vp, na.rm = T) - z*sd(FCO2vp, na.rm = T)/sqrt(100000)
}

write.csv(binned.nls.sf, "binned.nls.sf.csv")

```



